Pharmacologically inducing regenerative cardiac cells by small molecule drugs

Version 1

Abstract

Summary

In contrast to lower organisms, adult mammals cannot regenerate damaged hearts through de-differentiation of cardiomyocytes (CMs) into cells with regenerative capacity. Development of an enabling condition to induce such naturally unavailable cells with potential to proliferate and differentiate into CMs, i.e., regenerative cardiac cells (RCCs), in mammals will provide new insights and tools for heart regeneration. Here, a two-compound combination (2C), CHIR99021 and A-485, was identified to robustly induce RCCs from human embryonic stem cell (hESC) derived TNNT2+ CMs in vitro, which was confirmed by lineage tracing experiments. Functional analyses revealed that RCCs expressed a spectrum of genes essential for cardiogenesis and exhibited potential to become functional CMs, endothelial cells (ECs) and smooth muscle cells (SMCs). Consistent with the results in human cells, 2C-induced generation of RCCs were also observed in neonatal rats CMs in vitro and in vivo. Remarkably, administration of 2C can induce RCCs in adult mouse hearts and significantly improve survival and heart function in the mice undergoing myocardial infarction. Mechanistically, CHIR99021 is indispensable for transcriptional and epigenetic activation of genes essential for RCC, whereas A-485 mainly function to epigenetically down-regulate H3K27Ac and particularly H3K9Ac in CMs. Their combination specifically enhances both H3K27Ac and H3K9Ac on RCC genes, facilitating the establishment of RCC state dedifferentiated from CMs. Therefore, our findings demonstrated the feasibility and revealed the mechanisms of pharmacological induction of RCCs from endogenous CMs, which could offer a promising regenerative strategy to repair injured hearts.

Main Text

Lower organisms, such as zebrafish, possesses a robust cardiac regeneration capacity through CM dedifferentiation characterized by reactivation of embryonic cardiogenic genes and disassembly of their sarcomeric structures (Kikuchi & Poss, 2010; Lepilina et al, 2006). In contrast, limited cardiac regeneration was only permitted from postnatal day 1 to day 6 (PN1 to PN6) in mice with sarcomeric disassembly and proliferation of CMs in the resected section of heart (Porrello et al, 2011), which is no longer observed beyond this narrow time window. Therefore, discovery of small molecule drugs enabling induction of regenerative capacity in mammalian adult endogenous CMs by recapitulating the regenerative cellular state presented in zebrafish will shed mechanistic lights on the understanding of roadblocks preventing CMs from dedifferentiation in mammalian hearts, and facilitate development of regenerative medicine to treat various heart diseases caused by CM loss.

During mammalian heart development, cardiomyocytes (CMs) mainly originate from first heart field (FHF) and second heart field (SHF) (Srivastava, 2006). In particular, SHF cells marked by the pioneer transcription factor ISL1 can proliferate and differentiate into three major cardiovascular cell types, CMs, SMCs and ECs during embryonic heart development (Bu et al, 2009; Moretti et al, 2006). Consistently, a small number of ISL1+ cells residing in the neonatal mouse heart between postnatal Day 1 and Day 5 were reported to expand and differentiate into CMs under appropriate in vitro conditions (Laugwitz, 2005). Since ISL1 transcriptionally governs the expression of a cohort of cardiac genes, e.g., NKX2-5 (Ma et al, 2008), FGF10 (Watanabe et al, 2012), essential for embryonic cardiogenesis, it is reasonable to hypothesize that induced expression of ISL1 in CM may be a determinative step in the induction of regenerative cardiac cells (RCCs).

Toward induction of regeneration capacity in CMs, we conducted a screening for small molecules that can induce the expression of ISL1 in CMs. Our combinatorial screening identified a novel combination of CHIR99021 and A-485/2C that could unprecedentedly and efficiently induce dedifferentiation of human CMs into RCCs. These RCCs exhibited disassembled sarcomeric structures, high expression of embryonic cardiogenic genes, and increased number of CMs through re-differentiation in vitro. Further studies showed that 2C could robustly generate RCCs in hearts of adult mice and improve cardiac function in the mice with myocardial infarction. Our proof-of-concept discovery demonstrated that a simple combination of small molecule drugs can grant regenerative capacity to the heart by reprogramming CMs into RCCs.

2C treatment efficiently induces dedifferentiation of hESC-derived CMs

To obtain an adequate number of CMs for small molecule screening, we differentiated hESCs into CMs (SI Appendix, Fig. S1A and Movie S1) following a well-established step-wise protocol (Lian et al, 2013). Consistent with previous reports, nearly homogenous contracting CMs were observed on day 10 of differentiation. After purification with glucose-depleted medium containing abundant lactate (SI Appendix, Fig. S1B), highly pure TNNT2+ CMs were obtained, and then subsequently dissociated and seeded into 96-well plates. When CMs resumed contraction, they were treated with individual small molecules from a collection of over 4,000 compounds for 3 days (SI Appendix, Fig. S1C and Table S1), and then fixed and immunostained with ISL1. Using a high-content imaging (HCI) and analysis system, five compounds were initially identified to potentially induce regenerative capacity in CMs, indicated by induced sarcomere disassembly and ISL1 expression (SI Appendix, Fig. S1D). After further comparing the effects from various combinations, the unique combination of CHIR99021 and A-485 (2C) was found to most efficiently induce ISL1 expression with healthy cell state (SI Appendix, Fig. S1E).

When CMs with well-organized sarcomeres were treated with 2C in vitro, cells started to exhibit dedifferentiation-associated phenotype, such as reduction of cell size after 24 hours and growing as clusters after 48 hours and forming colonies after 60 hours (Fig. 1A and SI Appendix, Fig. S2 A-C). During this time, the expression of TNNT2 and MYL2 were gradually downregulated while ISL1 expression and the percentage of ISL1+ cells were gradually increased (SI Appendix, Fig. S2 D-G). Compared to untreated CMs, significant sarcomere disassembly and reduction of cytoplasmic/nuclear area were observed in cells treated with 2C for 60 hours (Fig. 1 B-D). In parallel, the number and percentage of TNNT2+ cells remained constant while ISL1+ cells increased remarkably (Fig. 1 E-G), accompanied by a ∼3-fold increase ISL1 expression at both mRNA and protein levels (Fig. 1 H-J). Of note, other typical genes involved in early embryonic cardiogenesis, including MESP1, LEF1 (Klaus et al, 2007), FUT4 (Wang et al, 2019), and NR2F2 (Churko et al, 2018) were also induced by 2C to express at high levels (Fig. 1J and SI Appendix, Fig. S2H). In addition, treatment with 2C resulted in significantly decreased expression of CM-specific genes, such as TNNT2, MYL2, MYL7, and MEF2C, while the pan-cardiac transcription factors, GATA4, TBX5, and NKX2-5, showed no obvious expression changes (Fig. 1J and SI Appendix, Fig. S2H). Furthermore, when we generated mature CMs from hESCs by treating differentiating cells with ZLN005 following a previously reported method (Liu et al, 2020), 2C could induce ISL1 expression in these mature CMs (SI Appendix, Fig. S2I). These results suggested that CMs might be converted by 2C into a state with regenerative capacity.

2C treatment induced dedifferentiation of hESC-derived CMs toward ISL1-expressing cells.

Cells induced from hESC-derived CMs by treatment with DMSO (NC) or 2C for 60 hours. A, Phase contrast images showing cell morphology. B, Immunofluorescence staining of the ISL1 (ISL1, green), and the CM marker cardiac troponin T (TNNT2, red) in the cells. Nuclei were stained by DAPI (4′,6-diamidino-2-phenylindole) and presented in DNA blue. (C and D), Cytosolic (C) and nuclear (D) areas of the cells. E, The number of total cells, TNNT2+ cells, and ISL1+ cells. F, Cell percentage of TNNT2+ cells. G, Fraction of ISL1+ cell in TNNT2+ cells. Error bars represent SD. ns, not significant (P > 0.05); **P < 0.01; ***P < 0.001. (H and I) Western blot (H) and quantitative analysis (I) of ISL1 expression in DMSO (NC) or 2C-treated CMs for 60 hours. Error bars represent SD. *P < 0.05; J, Heatmap illustration showing the fold-changes of indicted marker genes’ expression by 2C treatment, which were measured using qRT-PCR.

2C induced cells possess regenerative capacity

To confirm the regenerative capacity of 2C-induced cardiac cells, we investigated cell proliferation using a BrdU incorporation assay. Immunostaining clearly showed the co-localization of ISL1 and BrdU labeling in cells (Fig. 2A), and subsequent statistical analysis revealed a significant decrease in the nuclear area and increase in the cell percentage of ISL1+/BrdU+ positive cells following 2C treatment (Fig. 2 B and C). We next examined the potential for re-differentiation of 2C-induced cardiac cells into CMs. As expected, upon withdrawal of 2C and subsequent culture in CM media for 3 days (60h+3d), we observed the emergence of spontaneously contracting cells exhibiting CM-specific morphology with typical size of cytoplasmic/nuclear area (Fig. 2 D-F and Movie S2), downregulated expression of early embryonic cardiogenesis genes and upregulated expression of CM-specific genes, and the expression of pan-cardiac genes didn’t show any noticeable changes throughout this process (SI Appendix, Fig. S3A). Remarkably, there was a 1.4-fold increase in the number of re-differentiated CMs with clear TNNT2 staining relative to the initial CMs before 2C treatment (Fig. 2 G and H), demonstrating a regenerative potential in 2C-induced cardiac cells. In addition to CMs, 2C-induced cardiac cells were capable of differentiation towards SMA+ SMCs or CD31+/VE-Cadherin+ ECs in the presence of PDGF-BB and TGF-β1 or VEGF, bFGF and BMP4, respectively (Fig. 2I and SI Appendix, Fig. S3 B-D). In contrast, the expression of those genes was not detectable in DMSO-treated cells. Collectively, our findings indicate that treatment with 2C facilitates the conversion of CMs into cardiac cells with regenerative capability, including proliferative potential and ability to differentiate into the three major cardiovascular cell types, which were therefore named as regenerative cardiac cells (RCCs).

Regenerative ability of 2C-induced RCCs.

(A-C), Immunofluorescence staining (A) and statistical analysis (B and C) of the ISL1 (green) and BrdU (red) double positive RCCs induced from CMs by treatment with DMSO (NC) or 2C for 60 hours. DAPI (4′,6-diamidino-2-phenylindole) staining labeled nuclei as blue. Error bars represent SD. **P < 0.01. D, Phase contrast images of hESC-derived CMs treated by DMSO (NC) or 2C for 60 hours (60h) and subsequently cultured in the absence of 2C for another 3 days (60h+3d). (E and F), Cytosolic (E) and nuclear (F) areas of the cells under the same condition in (D). Error bars indicate SD. ns, not significant (P > 0.05); ***P < 0.001. G, Immunostaining showed the expression of ISL1 (green) and TNNT2 (red) in the cells under the same condition in (D). H, The fold change of TNNT2+ cells under the same condition in (D). Error bars represent SD. ns, not significant (P > 0.05); **P < 0.01. Immunostaining showed the expression of EC markers (CD31, green and CD144, red), SMC marker (SMA, red), and CM marker (TNNT2, green).

Lineage tracing demonstrated that 2C induced RCCs dedifferentiated from TNNT2+ CMs

Although TNNT2+ CMs purified by lactate-based culture medium were almost homogeneous populations, we noticed that these were still a low percentage (<4%) of purified cells still expressed ISL1 (SI Appendix, Fig. S1B). Therefore, we performed lineage-tracing experiments to further confirm that the 2C-induced RCCs were actually dedifferentiated from TNNT2+ CMs rather than proliferation of those residual ISL1+ cells. Upon directed cardiac differentiation of an ISL1mCherry/+ H9 hESC (K9) line established through CRISPR-based knock-in (SI Appendix, Fig. S4 A-D), mCherry expression faithfully reflected endogenous ISL1 expression (Fig. 3 A-C). After purification of K9-derived CMs, mCherry-negative cells were selected by FACS (Fig. 3 D and E), all of which were confirmed to be TNNT2+ CMs (SI Appendix, Fig. S4E). Those mCherry-negative CMs were then treated with 2C for 60 hours. The expression of TNNT2 and MYL2 were dramatically down-regulated, whereas the expression of ISL1 and mCherry was significantly up-regulated (Fig. 3 F and K). We found the induced mCherry+/ISL1+ RCCs exhibited significant morphological changes accompanied by a dramatic decrease in TNNT2 expression (Fig. 3F). In contrast, the mCherry-negative CMs remained high expression of TNNT2 with the typical sarcomere structure (Fig. 3F). FACS analysis further confirmed that 2C was capable of K9-derived mCherry-negative CMs into mCherry+/ISL1+ cells (Fig. 3 G and H). Consistently, similar results were observed using HUES7 hESC line with ISL1-mCherry knock-in reporter (K7) (Fig. 3 I, J, L, and SI Appendix, Fig. S5).

Lineage tracing demonstrated 2C induced dedifferentiation of TNNT2+ CMs to ISL1-expressing RCCs.

A, Immunofluorescence images showing expression of endogenous ISL1 (green) and ISL1-mCherry (red) reporter in the cells differentiated from K9 hESC KI reporter line at day 6 (D6). (B and C) Flow cytometry analysis of the percentage of mCherry+/ISL1+ cells in the cells differentiated from K9 at D6. (D and E) Flow cytometry analysis of the percentage of mCherry-negative cells at selection day 4 (SD4) in lactate purification medium. Error bars represent SD. (F-H) Cells induced from mCherry-negative CMs by treatment with or without 2C for 60 hours. Images showing the expression of mCHERRY (red) and TNNT2 (green) in the cells (F) and flow cytometry analysis of the percentage of mCherry-positive cells (G and H). Error bars represent SD. (I and J) Flow cytometry analysis of the percentage of mCherry-positive cells dedifferentiated from K7 hESC KI reporter line derived mCherry-negative CMs. Error bars represent SD. (K and L), Relative gene expression of ISL1, mCHERRY, LEF1, TNNT2 and MYL2 in K9- and K7-derived mCherry-negative CMs treated with DMSO (NC) or 2C for 60 hours, respectively. Error bars represent SD. *P < 0.05; **P < 0.01; ***P < 0.001. (M and N) Flow cytometric plots showing EGFP-labeled CMs by lineage-tracing of K9-derived mCherry-negative CMs (M), and bar graph showing the percentage of mCherry-negative CMs expressing EGFP (N). Error bars represent SD. O, Images showing the expression of ISL1 (red) and EGFP (green) in the cells induced from EGFP-positive/mCherry-negative CMs in (M) by treatment with or without 2C for 60 hours.

Furthermore, we used a lineage-tracing system to unequivocally examine the conversion of CMs to ISL1-expressing RCCs by 2C treatment. In this assay, CMs were transfected with plasmids encoding CreERT2 under the control of TNNT2 promoter and an EGFP reporter following flox-stop-flox cassette (SI Appendix, Fig. S6A). After optimizing virus titer and infection time (SI Appendix, Fig. S6 B-D), ∼0.6 of K9-derived mCherry-negative CMs were permanently labeled with EGFP following 6 days of treatment with tamoxifen (Fig. 3 M and N). Then, those EGFP-labeled mCherry-negative CMs were sorted by flow cytometry and subsequently treated with 2C for 60 hours. As expected, the expression of ISL1 was observed in these EGFP+ CMs (Fig. 3O). Collectively, these results robustly demonstrated that RCCs with ISL1 expression were truly generated by 2C treatment from TNNT2+ CMs.

2C-induced dedifferentiation of CMs provided a protective effect on cardiac infarction

In order to examine 2C’s reprogramming effect on CMs in vivo, we first tested 2C on primary neonatal rat CMs (Sakurai et al, 2014), on which 2C was found to consistently induce RCCs generation with corresponding ISL1 induction and morphological changes (SI Appendix, Fig. S7A). Then, we sought to examine whether 2C was capable of reprograming endogenous CMs to RCCs in vivo. For this purpose, neonatal SD rats were intraperitoneally administrated with 20 mg/kg of CHIR99021 and 10 mg/kg of A-485 daily for 5 days (SI Appendix, Fig. S7B). Compared to the vehicle/DMSO controls (NC), ISL1 was robustly induced in the cells from hearts of rats treated with 2C without apparent effects on body weight (SI Appendix, Fig. S7 C and D). These induced ISL1-expressing cells were widely distributed within the region 600 µm to 3000 µm down from the base of the hearts, including aorta (Ao), left/right atria, and both ventricles (SI Appendix, Fig. S7E). Notably, all of these ISL1+ expressed TNNT2 as well, strongly indicating 2C-induced RCCs were actually originated from CMs in vivo. More importantly, neither administration of 20 mg/kg CHIR99021 nor 10 mg/kg A-485 alone observed ISL1+ cells in ventricles, powerfully illustrating the combined effect of 2C on the induction of RCCs (SI Appendix, Fig. S8).

Similarly, RCCs were also efficiently induced in the Ao root and RA regions of the heart dissected from adult 129SvJ mice administrated with 2C (20 mg/kg of CHIR99021 and 10 mg/kg of A-485) for 5 consecutive days as well (Fig. 4A). Intriguingly, we sought to explore whether 2C-induced RCCs could rescue cardiac function in mice subjected to myocardial infarction (MI). In a prophylactic setting, mice were pre-treated with either 2C or vehicle/DMSO for 5 days (Fig. 4B) and then inducing MI by ligation of left anterior descending (LAD) artery. When measuring cardiac function with magnetic resonance imaging (MRI) on days 1, 8, 25 and 35 post-MI, we found pretreatment with 2C was able to significantly improve both cardiac function and survival rate in the mice, without affecting their body weight (Fig. 4 C-E). Next, when we treated mice with 2C or vehicle/DMSO post-MI in a therapeutic setting (Fig. 4F), remarkably mice with 2C treatment displayed a recovered cardiac function as assessed by both left ventricular ejection fraction (LVEF) and echocardiography (Fig. 4 G and H). Consistently, cardiac fibrosis was largely ameliorated in 2C-treated mice that displayed significantly smaller scar size in the hearts compared to control mice (Fig. 4I). Those in vivo studies collectively indicate that 2C or further improved small molecule drugs with similar mechanisms may have therapeutic utility to repair or regenerate hearts after cardiac injury. Altogether, these in vivo studies indicate that 2C-induced RCCs possess the same regenerative capacities comparable to those generated in vitro, which will shed light on development of small molecule drugs with therapeutic utility to repair or regenerate hearts after cardiac injury.

Heart regeneration via 2C-induced dedifferentiation of CMs.

A, Immunofluorescence staining of ISL1 (green) and TNNT2 (red) in cross-sectioned hearts from 2C or vehicle-treated (NC) adult 129SvJ mice. Ao, aorta. PA, pulmonary artery. LA, left atrial. RA, right atrial. IAS, interatrial septum. B, Schematic illustration of the method used to examine the prophylactic effect of 2C in 129SvJ mice post MI. C, Body weight of mice pre-treated with vehicle (DMSO) or 2C as shown in (B) at Day 12, Day 16, Day20, and Day35 after MI (1dpi). Error bars represent SD. ns, not significant (P > 0.05). D, Survival curve of sham-operated mice and mice pre-treated with vehicle (DMSO) or 2C as shown in (B), at indicated time points before or after MI. E, Ejection fraction (EF) of sham-operated mice and mice pre-treated with vehicle (DMSO) or 2C as shown in (B), before MI (baseline) or at Day 1, Day8, Day25, and Day35 after MI (1dpi). Error bars represent SD. ns, not significant (P > 0.05); **P < 0.01; ***P < 0.001. F, Schematic illustration of the method used to examine therapeutic effect of 2C in the 129SvJ mice post MI. G, Echocardiography of sham-operated mice and mice treated with vehicle (DMSO) or 2C as shown in (F) at Day 35 post MI. H, Serial fMRI measurements showing the cardiac function from sham-operated mice and mice treated with vehicle (DMSO) or 2C at as shown in (F), at Day 35 post MI. Error bars represent SD. *P < 0.05. I, Masson staining of serial transverse sections of hearts from sham-operated mice and mice treated with vehicle (DMSO) or 2C as shown in (F), at Day 35 post MI.

Both CHIR99021 and A-485 are required to induce dedifferentiation of CMs to RCCs

To gain insights into mechanisms involved in 2C-induced dedifferentiation of CMs to RCCs, we performed bulk RNA-seq analysis of K9-derived ISL1/mCherry-negative CMs treated with DMSO (NC) or 2C for 60 hours (Fig. 5A). Analysis of differentially expressed genes (DEGs) revealed a remarkable change in the transcriptome upon 2C treatment (Fig. 5B). In detail, 2C treatment led to increased expression of a large set of genes known to be associated with embryonic cardiogenic genes (e.g., MSX1, BMP4, TCF4, and LEF1) and suppression of a significant list of genes associated with CMs (Fig. 5C). Consistently, gene ontology (GO) analysis further revealed that down-regulated genes were mainly involved in cardiac maturation and muscle contraction (Fig. 5D), while up-regulated genes were enriched in cell junction and catenin complex (Fig. 5 E and F). Those embryonic cardiogenic genes and other genes significantly up-regulated (e.g., MSX2, NKD1, PDGFC, CTNNA2) after 2C treatment were confirmed by qPCR in K9-derived mCherry-negative CMs (Fig. 5G).

Bulk RNA-seq of analysis of 2C-treated ISL1/mCherry-negative CMs.

A, Scheme of bulk RNA-seq analysis of K9-derived mCherry-negative CMs with DMSO (NC) or 2C treatment for 60 hours. B, Heatmap of differentially expressed genes (DEGs) in ISL1/mCherry-negative CMs treated with DMSO (NC) or 2C for 60 hours. C, Volcano plot showing genes significantly changed by 60 hours of 2C treatment. (D and E) Gene Ontology (GO) analysis of downregulated (D) and upregulated (E) genes in ISL1/mCherry-negative CMs by 2C treatment for 60 hours, compared to DMSO (NC) treated cells. F, Plotting GO terms of upregulated genes by 2C treatment with cnetlpot. G, Relative expression fold-changes of indicated genes in K9-derived ISL1/mCherry-negative CMs by 60 hours of DMSO (NC) or 2C treatment. Error bars indicate SD. **P < 0.01; ***P < 0.001.

To investigate the individual effects of CHIR99021 and A-485 on induction of RCCs, we compared H9 hESC-derived CMs treated with CHIR99021 and A-485 individually or in combination. When measuring cytoplasmic/nuclear area, we found CHIR99021 significantly reduced both cytoplasmic and nuclear area, a distinguishable morphological feature of RCCs, in comparison to A-485 that only reduced cytoplasmic area (SI Appendix, Fig. S9 A-C). Consistently, disassembly of sarcomere structures was more frequently observed in CMs treated with CHIR99021 than in CMs treated with A-485 (SI Appendix, Fig. S9E). Although reduced expression of TNNT2 and MYL2 as well as increased expression of ISL1 can be induced individually by CHIR99021 and A-458 (SI Appendix, Fig. S9 D and E), CHIR99021 was able to uniquely promote the expression of a cohort of genes involved in embryonic cardiogenesis, including BMP4, NKD1, MSX2, PDGFC, LEF1 and TCF4 (SI Appendix, Fig. S9D). Upon withdrawal of CHIR99021, A-485 or 2C, the cell percentage of ISL1+ cell was similarly and significantly reduced (SI Appendix, Fig. S9F). Of note, a drastic increase in the number of TNNT2+ CMs was only observed in redifferentiation of 2C-treated cells, as confirmed by statistical analysis (SI Appendix, Fig. S9G). These findings suggested that CHIR99021 played a leading role in 2C-induced de-differentiation of CMs to RCCs, A-485 is also indispensable.

Reprogramming of CMs to RCCs by 2C went through an intermediate cell state

Next, we performed scRNA-seq of K9-derived mCherry-negative CMs treated with DMSO (NC) or 2C for 60 hours. Utilizing UMAP analysis, we observed the proportion of cells within three (cluster 0, 2, 3) out of 7 identified clusters increased after 2C treatment (Fig. 6 A and B). Among these three clusters, cells in cluster 2 highly expressed RCC characteristic genes, including those essential for embryonic cardiogenesis (e.g., ISL1, BMP4 and FGF20 (a key gene involved in expansion of early embryonic progenitor cells (Cohen et al, 2007)), and cell proliferation marker gene MKI67 (Fig. 6 C and D). By contrast, cells in cluster 0, 3, 4, and 5 expressed CMs specific markers (e.g., MYH6 and MYL2) at high levels (Fig. 6 C and D), indicating they were CMs. Meanwhile, cells in cluster 1 and 6 were considered as intermediate cells (ICs) during 2C-induced reprogramming of CMs to RCCs, because of expression of ACTA2, a marker gene reported to express in CM dedifferentiation, and genes associated with CM development, including COL1A1, COL1A2, and COL3A1 (Cui et al, 2020; Mononen et al, 2020) (Fig. 6 C and D).

Single-cell RNA-seq of 2C-treated mCherry-negative CMs.

A, UMAP analysis showing 7 clusters in cells induced from K9-derived mCherry-negative CMs by treatment with DMSO (NC) and 2C for 60 hours. B, The percentage of cells in the 7 indicated clusters, following DMSO (NC) or 2C treatment. C, Heatmap showing the differentially expressed genes in the cells from 7 indicated clusters. The representative marker genes of 7 indicated clusters were listed on the right. D, Violin plots showing the expression levels of marker genes of CMs (MYL2, MYH6), ICs (COL1A1, ACTA2), and RCCs (ISL1, BMP4, FGF20) among cells from 7 indicated clusters. E, UMAP analysis showing the second-level clustering of cluster 2 into 4 subclusters (left), which exhibited dramatic distinction under condition of 2C or NC (right). F, The percentage of cells in the 4 indicated subclusters within cluster 2, following DMSO (NC) or 2C treatment. G, Heatmap showing the differentially expressed genes among cells from 4 subclusters of cluster 2. Genes related to RCCs are highlighted in the green blocks on the right. (H and I) Pseudotime trajectory showing changes across various cell states upon 2C treatment, which were presented with different developmental pseudotime points (H) and cell states (I). J, Curves showing the dynamic expression of representative genes of RCCs (ISL1, FGF20), ICs (COL1A1), and CMs (TNNT2) along indicated pseudotime points.

When further examining cluster 2 with 435 cells from DMSO-treated cells and 410 cells from 2C-treated cells by second-level clustering analysis, we identified subcluster 0 (408, or 48.3% out of 845 cells) in cluster 2 which was exclusively observed in 2C-treated cells (Fig. 6 E and F). Furthermore, cells in subcluster 0 expressed a series of RCC-related genes essential for embryonic cardiogenesis, including MSX1, LEF1, BMP4, MSX2, and HAND1 (Fig. 6G), as well as a series of genes co-expressed with ISL1 in SHF progenitors in vivo, such as NR2F2, TBX5, ALCAM (Ghazizadeh et al, 2018), and CXCR4 (Andersen et al, 2018) (Fig. 6G), strongly indicating cells in subcluster 0 represent RCCs induced by 2C treatment. Thus, a unique gene set that included LIX1, NKD1, PDGFC, ARL4A, AXIN2, FZD7, ISL1, MSX1, MSX2, BMP4, LEF1, and HAND1 was found to determine RCC state during the reprogramming of CMs by 2C, and hence designated as RCC genes. In addition, the pseudo-time analysis that clearly revealed 2C-induced reprogramming of CMs (highly expressing TNNT2) went through ICs (highly expressing COL1A1), and ultimately towards RCCs (highly expressing ISL1 and FGF20) (Fig. 6 H-I).

RCC and CM genes were epigenetically regulated by 2C during reprogramming of CMs into RCCs

As a p300 acetyltransferase inhibitor, A-485 was previously reported to specifically inhibit acetylation of H3 at lysine 27 (H3K27Ac) but not at lysine 9 (H3K9Ac) in the PC3 cells (Lasko et al, 2017). Consistently, the percentage of H3K27Ac+ cells but not that of H3K9Ac+ cells were specifically and significantly reduced in CMs were treated with A-485 alone (SI Appendix, Fig. S10). In contrast, neither CHIR99021 nor 2C treatment resulted in noticeable alterations in the percentage of H3K27Ac+ and H3K9Ac+ cells (SI Appendix, Fig. S10). Intriguingly, further analyses revealed a remarkable reduction in the levels of H3K27Ac and H3K9Ac in TNNT2+ cells following A-485 treatment (SI Appendix, Fig. S10 F and L), suggesting that A-485 simultaneously inhibit both H3K27Ac and H3K9Ac on CM-specific genes during the dedifferentiation of CMs.

To gain more insights on the mechanism underlying 2C-induced silencing of CM genes and activation of RCC genes during reprogramming of CMs to RCCs, we performed chromatin immunoprecipitation-sequencing (ChIP-seq) analyses of H3K27Ac and H3K9Ac in DMSO, A-485, CHIR99021 and 2C-treated H9 hESC-derived CMs (Fig. 7A). We identified active promoters as proximal regions (±4 kb from the transcription start sites, or TSS) exhibiting H3K4me3 enrichment in the cells treated with DMSO or 2C (Fig. 7 B, and C). Treatment with 2C led to a striking increase in H3K9Ac and H3K27Ac enrichment around TSS regions on a whole-genome-wide scale (Fig. 7 B and C). Consistently, 2C treatment induced a significantly enhanced enrichment of H3K9Ac in 4,485 genes and H3K27Ac enrichment in 2,560 genes, whereas a decreased of H3K9Ac and H3K27Ac enrichment was detected in 346 and 1,834 genes respectively (SI Appendix, Fig. S11 A-F). Of note, a significant part of genes decreased H3K9Ac (268, or 58.1% out of 461) and H3K27Ac (777, or 68.6% out of 1132) enrichment in A485-treated cells were also observed among genes that H3K9Ac (268, or 77.5% out of 346) and H3K27Ac (777, or 42.4% out of 1834) enrichment following 2C treatment. Further GO analysis revealed that those genes decreased H3K9Ac and H3K27Ac enrichment in their promoters commonly seen in repose to A-485 or 2C treatment, preferentially exhibited cardiac specific functions, such as cardiac muscle contraction, sarcomere organization, and heart contraction (SI Appendix, Fig. S11 G-L). Compared to individual treatment with A-485 or CHIR99021, 2C treatment resulted in an increased H3K9Ac and H3K27Ac enrichment in the genes associated with cell cycle and cell division (SI Appendix, Fig. S11 K and L), strongly supporting the combined effects of 2C on RCC cell state establishment.

Chromatin immunoprecipitation-sequencing (ChIP-seq) analyses of chemical-treated H9 hESC-derived CMs.

A, Schematic illustration of ChIP-seq analysis of H9-derived CMs subjected to DMSO, A-485, CHIR99021, or 2C treatment for 60 hours. B, Average ChIP-seq signal profiles showing the indicated histone modifications around the TSS in the input and ChIP samples prepared from DMSO and 2C-treated cells. C, Heatmap showing the whole-genome wide distribution of H3K9Ac, H3K27Ac, and H3K4me3 peaks within a range of ±4kb from TSSs in the cells treated with DMSO or 2C for 60 hours. D, H3K9Ac and H3K27Ac peaks surrounding CM genes (TNNT2 and TNNI1) and RCC genes (LEF1 and ISL1) in the cells treated with DMSO, A-485, CHIR99021 or 2C for 60 hours, and H3K4me3 peaks surrounding the same genes in the cells treated by DMSO or 2C. E, Veen diagram showing the number of annotated genes with H3K9Ac or H3K27Ac enrichment in the cells treated with DMSO, A-485, CHIR99021 or 2C for 60 hours. Red circles indicate the number of genes with unique H3K9Ac or H3K27Ac enrichment induced by 2C treatment; black circles indicate the number of genes with H3K9Ac or H3K27Ac enrichment unaffected by any chemical treatment. (G-J) The annotated genes with the most significant changes in H3K9Ac enrichment following treatment with DMSO (G) or 2C (I) were ranked by Log10LR and analyzed by GOs (H and J), respectively. K, ISL1 binding motifs identified from the cells treated with A-485 or 2C.

Next, we specifically examined H3K9Ac and H3K27Ac peaks in the promoters of critical CM genes and RCC genes in the cells treated with or without A-485, CHIR99021 or 2C. We observed that both A-485 and CHIR99021 effectively reduced the peaks of H3K9Ac and H3K27Ac in the promoters of CM genes, including TNNT2, TNNI1, MYL7, MYH6, and MYH7. These reductions were further intensified with 2C treatment (Fig. 7D and SI Appendix, Fig. S12A). Interestingly, increased peaks of H3K9Ac and H3K27Ac in promoters of the RCC genes, such as LEF1, AXIN2, BMP4, LIX1, MSX1, MSX2, and NKD1, was only observed in condition with CHIR99021 available (Fig. 7D and SI Appendix, Fig. S12B). Other RCC genes, including ISL1, ARL4A, FZD7, HAND1, and PDGFC, exhibited a significantly elevated peaks of H3K9Ac and H3K27Ac by treatment with 2C and much higher than those treated with CHIR99021 alone (Fig. 7D and SI Appendix, Fig. S12C). Therefore, while CHIR99021 was able to activate the expression of RCC genes and simultaneously repressed CM genes partially through changing H3K27Ac and H3K9Ac during 2C-induced CM dedifferentiation to RCC, A-485 can inhibit CM genes directly through reducing of H3K27Ac and H3K9Ac modifications in the promoters of CM-specific genes, thereby assisting CHIR99021 to establish RCC state.

In addition, we observed that different small molecules had varied impacts on distribution of H3K9Ac and H3K27Ac enrichment during 2C-induced dedifferentiation of CM into RCC. As indicated by Veen diagram, distributions of H3K9Ac and H3K27Ac enrichment significantly diverged among cells treated with DMSO, A-485, CHIR99021, and 2C (Fig. 7E and F). For genes with H3K9Ac enrichment, 2716 (or 46.1%) out of 5891 annotated genes were only enriched in the cells treated with 2C, while only 527 (or 4.9%) out of 10785 annotated genes with H3K27Ac enrichment was exclusively observed (Fig. 7E and F). Thus, the process of 2C-induced dedifferentiation of CM to RCC is accompanied by vigorous alterations in H3K9Ac, whereas H3K27Ac is remain relatively stable, indicating that the genome-wide distribution of H3K9Ac exhibited drastic distinction between CMs and RCCs. To further confirm this notion, we performed a comparative analysis of annotated genes between the cells treat with 2C or DMSO to validate the most significant differential genes (i.e., Log10LR>3) under the respective condition. In comparison to 430 genes with H3K9Ac enrichment observed in DMSO-treated cells, which were mainly involved in cardiac muscle contraction (Fig. 7G and H), 2C treatment markedly enhanced the enrichment of H3K9Ac in 123 genes associated with the transcription factor complex and Wnt signaling pathway (Fig. 7 I and J). In particular, 12 CM genes, including TNNT2, were ranked in top 90 out of 430 annotated genes with H3K9Ac enrichment. In contrast, 7 RCC genes, including ISL1, were ranked in the top 50 out of 123 annotated genes with H3K9Ac enrichment. Moreover, DNA-binding motif analysis indicated that the ISL1 binding motifs only exhibited accessibility in the cells treated with A-485 or 2C (Fig. 7K), underscoring the vital role of A-485 in 2C treatment to activate the expression of RCC genes. Taken together, these findings revealed that CHIR99021 is responsive for transcriptional and epigenetic activation of genes essential for cardiac embryonic development, while A-485 promotes conversion of CMs to RCCs by epigenetic downregulation of H3K27Ac and particularly H3K9Ac in CMs, which together facilitate the induction of RCCs from CMs.

Discussion

It is worth noting that the RCCs were only generated in hearts of 2C-treated postnatal rats, but not in rats after treatment with only 20 mg/kg CHIR99021 or 10 mg/kg A-485 alone (SI Appendix, Fig. S8), indicating the reprogramming of endogenous CMs to RCCs depends on the cooperative effect from these two molecules. CHIR99021 has been well characterized as a GSK3 inhibitor to activate the Wnt signaling pathway. Previous studies have elucidated the role of CHIR99021 in initiating proliferation of hPSC-derived CMs (Buikema et al, 2020; Quaife-Ryan et al, 2020), accompanied by induction of H3K27Ac on promoter regions of multiple cardiac lineage genes, including TNNT2, MYH6, and ISL1 (Quaife-Ryan et al., 2020). However, this proliferation was only achieved in neonatal CMs in vivo (Buikema et al., 2020; Quaife-Ryan et al., 2020). Consistently, we also found the enhanced levels of H3K27Ac in TNNT2+ CMs by treatment with CHIR99021 alone or 2C (Fig. 7D and SI Appendix, Fig. S10F). In addition, we found that the p300/CBP inhibitor A-485 facilitate the reprogramming of CMs into RCCs by epigenetically inhibiting expression of CM-specific genes such as TNNT2 (Fig. 7D and SI Appendix, Fig. S9D), while enhancing the modifications of H3K9Ac and H3K27Ac on RCC-specific genes such as ISL1 in combination with CHIR99021 (Fig. 7 D and K). These findings promoted us to postulate that A-485 and its mediated mechanism may be used to modulate the cell fate of other terminally differentiated cells by epigenetically suppressing the expression of corresponding lineages specific genes.

2C-induced RCCs exhibited potential to differentiate into three cardiac-lineages of the heart in vitro, which might contribute additional beneficial effects in heart repair and regeneration. Furthermore, RCCs expressed a broad set of genes essential for embryonic cardiogenesis (Fig. 1J and Fig. 5G), including ISL1, and a series of potentially important genes identified in this study (Fig. 6G), all of which were highly expressed in ISL1+ SHF cardiac progenitors during mammalian heart development. However, RCCs had a restricted proliferation capacity, although they were able to be passaged 3-5 times under 2C conditions (data not shown), which is distinct to ISL1+ cardiac progenitors (Bu et al., 2009; Moretti et al., 2006). These findings collectively indicated that 2C-induced RCCs dedifferentiated from CMs may be expandable under an appropriate condition and deserve further investigation.

Supporting information

eLife Assessment

This manuscript offers valuable information on the effect of two small molecule combinations (2C), CHIR99021 and A-485, during the reprogramming of mature cardiomyocytes into regenerative cardiac cells. This manuscript is incomplete, as the mechanistic insights derived from transcriptomic and genomic datasets are without experimental validation. This manuscript also needs additional experimental support to confirm the regenerative potential of 2C and improvements in the data analysis and presentation. Overall, this interesting work provides insights into the development of therapeutic targets for cardiac regeneration in infarcted hearts.

Reviewer #1 (Public Review):

Summary:

This manuscript reports that a combination of two small molecules, 2C (CHIR99027 and A-485) enabled to induce the dedifferentiation of hESC-derived cardiomyocytes (CMs) into regenerative cardiac cells (RCC). These RCCs had disassembled sarcomeric structures and elevated expression of embryonic cardiogenic genes such as ISL1, which exhibited proliferative potential and were able to differentiate into cardiomyocytes, endothelial cells, and smooth muscle cells. Lineage tracing further suggested that RCCs originated from TNNT2+ cells, not pre-existing ISL1+ cells. Furthermore, 2C treatment increased the numbers of RCC cells in neonatal rat and adult mouse hearts and improved cardiac function post-MI in adult mice. Mechanistically, bulk RNA-seq analysis revealed that 2C led to elevated expression of embryonic cardiogenic genes while down-regulation of CM-specific genes. Single-cell RNA-seq data showed that 2C promoted cardiomyocyte transition into an intermediate state that is marked with ACTA2 and COL1A1, which subsequently transformed into RCCs. Finally, ChIP-seq analysis demonstrated that CHIR99027 enhanced H3K9Ac and H3K27Ac modifications in embryonic cardiac genes, while A-485 inhibited these modifications in cardiac-specific genes. These combined alterations effectively induced the dedifferentiation of cardiomyocytes into RCCs.

Strengths:

Overall, this work is quite comprehensive and is logically and rigorously designed. The phenotypic and functional data on 2C are strong.

Weaknesses:

The mechanistic insights of 2C are primarily derived from transcriptomic and genomic datasets without experimental verification.

Reviewer #2 (Public Review):

Summary:

The ability of cardiac cells to regenerate has been the object of intense (and sometimes controversial) research in biology. While lower organisms can robustly undergo cardiac regeneration by reactivation of the embryonic cardiogenic pathway, this ability is strongly reduced in mice, both temporally and qualitatively. Finding a way to derive precursor cells with regenerative ability from differentiated cells in mammals has been challenging.

Zhou, He, and colleagues hypothesized that ISL-1-positive cells would show regenerative capacity and developed a small molecules screen to dedifferentiate cardiomyocytes (CM) to ISL1-positive precursor cells. Using hESC-derived CM, the authors found that the combination of both, WNT activation (CHIR99021) and p300 acetyltransferase inhibition (A-485) (named 2C protocol) induces CM dedifferentiation to regenerative cardiac cells (RCCs). RCCs are proliferative and re-express embryonic cardiogenic genes while decreasing the expression of more mature cardiac genes, bringing them towards a more precursor-like state. RCCs were able to differentiate into CM, smooth muscle cells, and endothelial cells, highlighting their multipotent property. In vivo, administration of 2C in rats and mice had protective effects on myocardial infarction.

Mechanistically, the authors report that the 2C protocol drives CM-specific transcriptional and epigenetic changes.

Strengths:

The authors made a great effort to validate their data using orthogonal ways, and several hESC lines. The use of lineage tracing convincingly showed a dedifferentiation from CM. They translate their findings into an in vivo model of myocardial injury, and show functional cardiac regeneration post-injury. They also showed that 2C could surprisingly be used as a preventive treatment. Together their data may suggest a regenerative effect of 2C both in vitro and in vivo settings. If confirmed, this study might unlock therapeutic strategy for cardiac regeneration.

Weaknesses:

Several points remain puzzling to me and some aspects of this study need to be clarified and extended:

General comments:

* Experimental design & Interpretation*

1. The main hypothesis (line 50) that Isl1 cells have regenerative properties is not extremely novel (10.1172/jci.insight.80920, doi.org/10.1038/nature03215,10.1016/s1534-5807(03)00363-0).

2. Based on Table S1, concentration of A-485 used in the screen is 10uM but used throughout this study at 0.5um. Could the authors provide a rationale for this 20x reduction of concentration? It would be useful to get a titration of this compound for the effects tested.

3. It is confusing to clearly understand what proportion of CMs dedifferentiate to become RCCs. The lineage tracing data suggests only 0.6%-1.5% of cells undergo this transition. It is difficult to understand how such a small fraction can have wide effects in their different experimental settings. This is specifically true when the author quantified nuclear and cytosolic area on brightfield pictures - would the same effect on nuclear/cytosolic area be observed in Isl1 KO cells?

4. The authors totally disregard the effect of i-BET-762 that gives a very similar percentage of Isl1-positive cells when combined to 2C (Supp. 1E). What is the effect of CHIR + I-BET-762 alone?

5. It is really hard to understand the contradictory effects of A-485 on acetylation status. The authors mentioned that A-485 only has an effect on H3K27Ac and not on H3K9Ac (line 221) to later (line 226) contradict themselves by saying it also has an effect on H3K9Ac. To explain this discrepancy, the authors vaguely mentioned "further analyses" without giving any other details. It would be transparent to explain what led to this radical change in interpretation.

6. The difference in the ChIP peak height is rather minimal for the H3K9Ac data. Were the peaks normalized to the sequencing depth? What does the y-axis represent on these ChIP traces (number of counts?)

7. Would it be possible to test this 2C protocol on mESC and see if similar changes occur? How transcriptionally different would these mouse RCCs be to Isl1+ progenitors isolated from neonatal mice (P1-P5)?

Statistics & Data acquisition

1. The authors mentioned experiments were done at least 3 times and each dot plotted on a graph is an average of technical repeat for one biological repeat. My understanding would be that if I see 9 dots, it means the experiment has been done 9 times - What would be the rationale for such a high number of repeats? It is an "artificial" way to increase the power of a test and might lead to misinterpretation of the data. This becomes relevant for some figures where biological difference is minimal and they still show statistical differences (e.g. Supp 2E, Supp 3A, Supp 9C,...). This is also true for in vivo section (Fig. 4G).

It would help to have a precise clarification between technical and biological repeats in the figure legends (e.g. n=3 biological repeat (aka 3 dots on a graph) obtained from averaging XX technical repeats), as well as the specific test stated the legend in addition to the general paragraph in the methods. Providing raw numerical data so readers can re-test them independently would also be a transparent way to do it.

1. Does the author test for normality before applying a specific test? Please clarify and justify either way.

2. If each dot represents a biological repeat as stated in the method section, why do some datasets have different numbers of repeats between NC and 2C if obtained in parallel? Have repeats been excluded?

Reviewer #3 (Public Review):

Summary:

The manuscript by Zhou and colleagues describes the potential of a two-compound combination (2C), CHIR99021 and A-485, which can generate regenerative cardiac cells (RCCs) from human embryonic stem cell-derived TNNT2+ cardiomyocytes. The authors have also demonstrated this phenomenon in neonatal rats CMs in vitro. Further, the administration of 2C can generate RCCs in adult mouse hearts and significantly improve survival and cardiac function in mice subjected to myocardial infarction. Interestingly, 2C treatment induces global changes in transcription and epigenetic modifications.

Strengths of the study:

1. This study describes the potential of 2C in improving the regeneration of the heart post-MI. The findings may have a translation potential. The idea of promoting the regenerative capacity of the heart by reprogramming CMs into RCCs is interesting.

2. The authors have validated the effect of 2C independently in hESCs, rat CMs, and a model of MI.

3. The authors explored the mechanism by Single-cell RNA-seq and Chip-Seq, which points to the transcriptional and epigenetic activation of genes essential for RCC.

Weaknesses of the study:

1. The mechanism involved in the 2C-mediated generation of RCCs is still unclear. The leads found in the RNA-seq and ChIP-seq were not validated experimentally.

2. Considering the very low number of RCCs (0.6%-1.5% of cells) generated, I cannot comprehend how the heart is protected from MI. Did the author believe 2C would affect the survival or metabolism of existing CM under hypoxia? What percentage of cells were regenerated by 2C treatment post-MI?

3. I would like to know about administering 2C in mice, which could have generated RCCs- dedifferentiated CMs in the heart. Does 2C affect the cardiac functions in mice under basal conditions? Also, does 2C administration affect any physiology in mice? The cardiac structural and functional parameters are required post-2C administration.

4. It is also not tested whether 2C would affect other cell types of the heart, including fibroblasts and endothelial cells, in vitro and in vivo. Assuming the level of protection by 2C in mice, it would affect other cell types.

5. It is still being determined how the authors chose the dose of 2C for in vivo and in vitro studies, although the concentration used for screening is different. Assessing the effect of 2C in a dose-dependent manner is essential.

6. A-485 affects H3K27Ac but not H3K9Ac. However, data show that both H3K27Ac and H3K9Ac are affected. An explanation is required.

7. The authors use "regeneration" even at the screening stage. I am wondering if regeneration could be assessed by the experimental approach they adopted.

Reviewer #4 (Public Review):

Overview:

The present manuscript by Zhou and colleagues investigates the impact of a new combination of compounds termed CHIR99021 and A-485 on stimulating cardiac cell regeneration. This manuscript fits the journal and addresses an important contribution to scientific knowledge. However, the following major revisions need to be addressed as stated below.

Major comments:

-The authors should include more information that clarifies and justifies their hypothesis.

-The story line is not well developed and thus not convincing since the results from different sections are not well connected.

-The main text needs to be improved, and authors should explain their purpose in choosing to study ISL1-CMs. Also, to well argument why they conducted this study and its significance.

-Page 3, row 57-58: Please add the references.

-Page 3-4, row 67-68, authors stated "When CMs resumed contraction, they were treated with individual small molecules from a collection of over 4,000 compounds for 3 days (SI Appendix, Fig. S1C and Table S1), and then fixed and immunostained with ISL1". Please explain better, and show the results of the selected screening compounds.

-Authors must make an effort to discuss their findings in a bold way in order to provide a comprehensive and articulate explanation of their results to the readers. There is much information missing from this section. This should also propose new research avenues and foresee the challenges in future investigations.

-Authors must include a conclusion and future perspectives of this study.

- Page 4, row 73, the authors stated that the unique compound combination 'CHIR99021 and A-485' was found to be the most efficient in promoting ISL1 expression with a healthy cell state. However, the authors should prove that by showing at least the cell viability of these compound combinations at different concentrations and timings as a supplementary figure.

-There is some missing information in the methods part, for example, "Images were captured using a confocal Zeiss LSM710 and Olympus IX83 inverted microscope"; authors should include the objective used and the image size, and should include which method they used to analyze the acquired images.

-Figure S3A shows that the TNNT2 mRNA expression was completely absent after 60 hours of 2C administration. Authors should explain this further.

-Figure 3J, there is high variability in the graph of mCherry cells (%). Please choose a better graph, or increase the independent experiment.

-Authors did not explain/discuss their results of the DNA-binding motif analysis of ISL1 in the cells treated with A-485 or 2C (Figure 7K).

-Figure S1B and D: the image's labeling is not clear. In the exact same figure S1B, how can the authors explain the reduction of ISL cells? Do the authors make the treatment with the compound CHIR99021 as shown in figure S1A? If so, the authors should clarify the ISL reduction in Figure S1B.

-Figure 1H: please improve the immunoblot, the level of B-actin does not match among the different conditions, or provide a relative quantification of the proteins.

-Please indicate further information in the methodology part about the compounds used in this study.

-Figures are not well justified and figure legends are not sufficient enough to explain the figures.

-Please improve the figure legends by including more further information; for example, in Figure 2SH it is highlighted only the "DAPI (4′,6-diamidino-2- phenylindole) staining labeled nuclei as blue" but how about the other markers?

-Figure 2F: the graph shows some high variations in "ns" between NC at 2C and in 60h+3d. I would recommend increasing the independent experiments. Similar observation goes also for figure 2E.

-Authors should provide limitations of this study.

Author response:

The following is the authors’ response to the original reviews.

Reviewer 1

Overall, this work is quite comprehensive and is logically and rigorously designed. The phenotypic and functional data on 2C are strong.

Thank you for your positive feedback on our findings!

(1) Comment from Reviewer 1 suggesting the mechanistic insights of 2C are primarily derived from transcriptomic and genomic datasets without experimental verification.

Thank you for emphasizing the importance of experimental validation to support our transcriptomic and genomic findings. We acknowledge the gap in direct experimental evidence for the mechanistic insights of section 2C and recognize the value of such validation in strengthening our conclusions. While we recognize the importance of such validation, our current dataset lacks the comprehensive preliminary results necessary for inclusion in the supplemental material. We believe that the mechanistic insights presented offer a substantial foundation for the future research, where we aim to explore these aspects in depth with targeted experimental approaches.

Reviewer 2

Together their data may suggest a regenerative effect of 2C both in vitro and in vivo settings. If confirmed, this study might unlock therapeutic strategy for cardiac regeneration.

Thank you for your positive comment on the significance of our findings and the valuable therapeutic potential of 2C in cardiac regeneration!

(1) Comment from Reviewer 2 pointing out the the main hypothesis (line 50) that Isl1 cells have regenerative properties is not extremely novel.

We agree with the reviewer that Isl1-positive cells possess regenerative properties. Following the reviewer’s suggestion, we have revised the original wording (line 46 in the revised manuscript).

(2) Comment from Reviewer 2 asking for providing a rationale for this 20x reduction of A-485 concentration? It would be useful to get a titration of this compound for the effects tested.

As suggested by the reviewer, we have added the titration results of A-485 in Figure 1—figure supplement 1F-G.

(3) Comment from Reviewer 2 confusing to clearly understand what proportion of CMs dedifferentiate to become RCCs. The lineage tracing data suggests only 0.6%-1.5% of cells undergo this transition. It is difficult to understand how such a small fraction can have wide effects in their different experimental settings. This is specifically true when the author quantified nuclear and cytosolic area on brightfield pictures - would the same effect on nuclear/cytosolic area be observed in Isl1 KO cells.

We appreciate the reviewer's insightful observation on the proportion of CMs undergoing dedifferentiation into RCCs and the potential impact of this subset on our experimental outcomes. The lineage tracing data indicating that only 0.6%-1.5% of CMs transition to RCCs indeed reflects a modest proportion. This observation raises valid questions regarding the broader implications of such a limited fraction in the context of cardiac regeneration and the experimental effects reported. It's important to note that while the proportion of CMs dedifferentiating into RCCs is small, the biological significance and potential impact of these RCCs could be disproportionately large. Emerging evidence suggests that even a minimal number of stem or progenitor cells can exert significant effects on tissue repair and regeneration, possibly through paracrine mechanisms or by acting as key signaling centers within the tissue microenvironment (Fernandes et al., 2015). Regarding the specific question about 2C’s effects on nuclear/cytosolic area in Isl1 knockout (KO) cells, we appreciate the suggestion and consider that such comparative studies would provide valuable insights for future comprehensively understanding the significant impact of 2C-induced RCCs in future search. In addition, ISL1 KO cells are also described in detail in the article published in eLife in 2018 by Quaranta et al.

(4) Comment from Reviewer 2 asking for the effect of CHIR + I-BET-762 alone.

As suggested by the reviewer, we have added the results of CHIR + T-BET-762 in Figure 1—figure supplement 1H.

(5) Comment from Reviewer 2 suggesting a transparent explaination about the effects of A-485 on acetylation status.

We thank the reviewer for highlighting the confusion regarding the effects of A-485 on the acetylation status of H3K27Ac and H3K9Ac. Upon re-examination of our data and statements, we recognize the need for clarity in our explanation and the inconsistency it may have caused (lines 223-231 on page 8).

Initially, our observations suggested a selective effect of A-485 on H3K27Ac based on early experimental results (Figure 7—figure supplement 1). This conclusion was drawn from preliminary analyses that focused predominantly on this specific histone mark. However, upon further comprehensive examination of our data, including additional replicates and more sensitive detection methods, we observed that A-485 also impacts H3K9Ac levels (Figure 7—figure supplement 1F). This latter finding emerged from expanded datasets that were not initially considered in our preliminary conclusions.

The "further analyses" mentioned referred to these subsequent experimental investigations, which included chromatin immunoprecipitation (ChIP) assays and extended sample sizes, providing a more robust dataset for evaluating the effects of A-485. We understand the importance of transparency and rigor in scientific communication. To address this, we have revised the manuscript to clearly delineate the progression of our analyses and the evidential basis for our revised understanding of A-485's effects. This includes a detailed description of the methodologies employed in our follow-up experiments (line 537 on page 27), the statistical approaches for data analysis (lines 226-227 in supporting information), and how these led to the updated interpretation regarding A-485's impact on histone acetylation (lines232-269).

(6) Comment from Reviewer 2 asking for the difference in the ChIP peaks representation of the y-axis on the ChIP traces.

Thank you for raising this quest. Actually, we did not normalise the sequencing depth and the y-axis represents the number of counts (line 537 on page 27 and lines 226-227 in supporting information).

(7) Comment from Reviewer 2 suggesting the possibility of testing this 2C protocol on mESCs to see if similar changes are subject to and how these mouse RCCs differ transcriptionally from Isl1+ progenitor cells isolated from neonatal mice (P1-P5)?

Thank you for your insightful questions. Testing the 2C protocol on mouse embryonic stem cells (mESCs) to observe if similar changes occur presents an excellent opportunity to further validate the versatility and applicability of our findings across different stem cell models. We agree that such experiments would not only strengthen the current study but also provide valuable insights into the conservation of mechanisms across species. We are currently in the process of setting up experiments to address this very question and anticipate that the results will significantly contribute to our understanding of cardiomyocyte differentiation processes. Regarding the transcriptional comparison between mouse regenerative cardiac cells (RCCs) induced by our 2C protocol and Isl1+ progenitors isolated from neonatal mice (P1-P5), this comparison is indeed crucial for delineating the specific identity and developmental potential of the RCCs generated. However, a comprehensive side-by-side transcriptomic analysis is required to systematically identify these differences and understand their biological implications. We plan to undertake this analysis as part of our future studies, which will include detailed RNA sequencing and comparative gene expression profiling to elucidate the transcriptional similarities and differences between these cell populations. These future directions will enhance our current findings, provide a deeper mechanistic understanding, and confirm the potential of the 2C protocol in regenerative medicine applications. We appreciate the reviewer's suggestions and acknowledge the importance of these experiments in advancing the field.

(8) Comment from Reviewer 2 with a suggestion to have a precise clarification of statistics & data acquisition.

As suggested by the reviewer, we have revised clarifications to make them clearer (lines 228-233 in supporting information and a precise description of each paragraph involving statistical analyses).

Reviewer 3

The findings may have a translation potential. The idea of promoting the regenerative capacity of the heart by reprogramming CMs into RCCs is interesting.

Thank you for your appreciation of the significance and translational potential of our findings!

(1) Comment from Reviewer 3 suggesting the mechanism involved in the 2C-mediated generation of RCCs is unclear and the lead found in the RAN-seq and ChIP-seq are not experimatally validated.

We acknowledge the reviewer's concern regarding the lack of experimental validation for the mechanisms identified through RNA-seq and ChIP-seq analyses in the generation of RCCs from the 2C state. We understand the importance of substantiating these molecular leads with empirical data to strengthen our conclusions. Currently, our findings are based on in-depth bioinformatic analyses, which have provided us with valuable insights and a strong basis for hypothesis generation. Moving forward, we plan to prioritize experimental validation of key pathways and targets identified in our study. This will include designing targeted experiments to elucidate the functional roles of these mechanisms in the 2C-mediated generation of RCCs. We appreciate the opportunity to clarify our approach and future directions, and we are committed to addressing this gap in subsequent work.

(2) Comment from Reviewer 3 considering the very low number of RCCs (0.6%-1.5% of cells) generated cannot protect the heart from MI, and whether 2C affects the the survival or metabolism of existing CM under hypoxia conditions, and what percentage of cells are regenerated by 2C treatment post-MI?

We appreciate the reviewer's insightful queries regarding the protective effects of 2C treatment against myocardial infarction (MI) given the low percentage of RCCs generated. It is our hypothesis that the benefits of 2C treatment extend beyond mere cell numbers. We propose that 2C may enhance the survival and metabolic resilience of existing CMs under hypoxic conditions, thereby contributing to cardiac protection post-MI. Our future investigations will aim to quantify the precise percentage of cells regenerated by 2C treatment post-MI and explore its broader impacts on cardiac tissue survival and repair mechanisms.

(3) Comment from Reviewer 3 suggesting the administration of 2C in mice, as well as whether 2C affects cardiac function under basal conditions and any physiology in mice, and the need to examine cardiac structural and functional parameters after administration of 2C.

We appreciate the reviewer's interest in the potential effects of 2C administration on cardiac function and overall physiology in mice. While we observed a decrease in body weight at P5 compared to controls, our immunofluorescence staining did not indicate any changes in cardiac structure (Figure 4— figure supplement 1E). This suggests that while 2C administration impacts neonatal rat physiology, it does not adversely affect cardiac structure under basal conditions. Further investigations are planned to assess the functional parameters of the heart post-2C administration to comprehensively understand its effects.

(4) Comment from Reviewer 3 suggesting the potential effects of 2C on other cell types of the heart, including fibroblasts and endothelial cells, in vitro and in vivo.

We value the reviewer's suggestion to explore the effects of 2C on various cardiac cell types, including fibroblasts and endothelial cells, both in vitro and in vivo. We acknowledge the importance of understanding the broader impact of 2C treatment across different cell populations within the heart, given its potential protective effects. To address this, we are designing a series of experiments to assess 2C's influence on these cell types, aiming to elucidate any changes in their behavior, proliferation, and function following treatment. This comprehensive approach will allow us to better understand the mechanistic basis of 2C's cardioprotective effects.

(5) Comment from Reviewer 3 suggesting validation the effect of 2C in a dose-dependent manner.

As suggested by the reviewer, we have supplemented the effect of 2C in dose-dependent (Figure 1— figure supplement 1F-G).

(6) Comment from Reviewer 3 suggesting an explanation of how A-485 affects H3K27Ac and H3K9Ac.

We appreciate the reviewer pointing out the discrepancy regarding the effects of A-485 on H3K27Ac and H3K9Ac. Upon re-examination of our data, we realize that our initial interpretation may have overlooked the broader impact of A-485 on histone acetylation patterns. It appears that A-485 does indeed influence both H3K27Ac and H3K9Ac, contrary to our initial statement. This oversight will be corrected in our revised manuscript, where we will provide a more detailed analysis and discussion of A-485's impact on these histone marks, alongside an explanation for the observed effects (lines 223-269 across page 8-9).

(7) Comment from Reviewer 3 with a correction to use "regeneration" at the screeing stage.

As suggested by the reviewer, we have amended the wording in the text (line 66 on page 3).

Reviewer 4

Comment from Reviewer 4 suggesting more information that clarifies and justifies the hypothesis.

As suggested by the reviewer, we added more information to clarify and justify the hypothesis (lines 39-47 on page 3).

(1) Comment from Reviewer 4 pointing out the story line is not well developed.

To address the reviewer’s question, we revised the manuscript to ensure a smooth and coherent logical flow.

(2) Comment from Reviewer 4 pointing out the purpose in choosing to study ISL1-CMs.

As raised by the reviewer, we have clarified the rationale for using ISL1 as a marker to define RCCs in revised manuscript (lines 39-47 on page 3).

(3) Comment from Reviewer 4 pointing out the missing references in row 57-58.

Thank you for pointing this out, we fixed it.

(4) Comment from Reviewer 4 suggesting more explains and show the results of the screening compounds.

As suggested by the reviewer, we added additional explanations in lines 65-73 and showed the screening results in Figure 1—figure supplement 1F-H.

(5) Comment from Reviewer 4 suggesting an in-depth discussion of the findings.

Thank you for the suggestion, we included additional discussion at the end of the article.

(6) Comment from Reviewer 4 suggesting a conclusion should be inculded in the main text.

Thank you for the suggestion, we made a revision.

(7) Comment from Reviewer 4 pointing out the cell viability under different concentrations of 2C.

As mentioned by the reviewer, have supplemented the cell numbers during different doses of 2C treatment (Figure 2F).

(8) Comment from Reviewer 4 pointing out the missing information in the methods.

Thank you for the suggestion, we made additions.

(9) Comment from Reviewer 4 suggesting more explanations in Figure S3A.

As mentioned by the reviewer, we made a revision in original Fig.S3A (now is Figure 2—figure supplement 1).

(10) Comment from Reviewer 4 pointing out the high variability of mCherry cells (%) in Figure 3J.

Thank you. We made a revision.

(11) Comment from Reviewer 4 suggesting more explanations on the DNA-binding motif of ISL1 in the cells treated with A-485 or 2C.

Thank you for the suggestion, we added additional explanations (lines 270-274 on page 9).

(12) Comment from Reviewer 4 pointing out the unclear labeling in Figure S1B and D.

Thank you for the suggestion, made a revision (lines 240-245 in supporting information).

(13) Comment from Reviewer 4 suggesting a relative quantification of the proteins in Figure 1H.

Thank you for the suggestion. We have quantified the relative expression levels of proteins in original Fig. 1H. As shown in Figure 1F.

(14) Comment from Reviewer 4 suggesting to provide detailed information in the methodology part about the compounds.

Thank you for the suggestion, we made a revision.

(15) Comment from Reviewer 4 pointing out the insufficient explanations on figure legends.

Thank you for the suggestion, we made a revision.

(16) Comment from Reviewer 4 suggesting more independent experiments to reduce the high variations in “ns” between NC and 2C at 60h+3d shown in Figure 2E and F.

Thank you for the suggestion, we made a revision in Figure 2F.

(17) Comment from Reviewer 4 suggesting a limitations should be provided in the text.

Thank you for the suggestion, we have made provide a limitation statement in the revised manuscript (lines 300-311 on page 10).

Version 2

Abstract

Summary

Adult mammals, unlike some lower organisms, lack the ability to regenerate damaged hearts through cardiomyocytes (CMs) dedifferentiation into cells with regenerative capacity. Developing conditions to induce such naturally unavailable cells with potential to proliferate and differentiate into CMs, i.e., regenerative cardiac cells (RCCs), in mammals will provide new insights and tools for heart regeneration research. In this study, we demonstrate that a two-compound combination, CHIR99021 and A-485 (2C), effectively induces RCCs from human embryonic stem cell (hESC)-derived TNNT2+ CMs in vitro, as evidenced by lineage tracing experiments. Functional analysis shows that these RCCs express a broad spectrum of cardiogenesis genes and have the potential to differentiate into functional CMs, endothelial cells (ECs), and smooth muscle cells (SMCs). Importantly, similar results were observed in neonatal rat CMs both in vitro and in vivo. Remarkably, administering 2C in adult mouse hearts significantly enhances survival and improves heart function post-myocardial infarction. Mechanistically, CHIR99021 is crucial for the transcriptional and epigenetic activation of genes essential for RCC development, while A-485 primarily suppresses H3K27Ac and particularly H3K9Ac in CMs. Their synergistic effect enhances these modifications on RCC genes, facilitating the transition from CMs to RCCs. Therefore, our findings demonstrate the feasibility and reveal the mechanisms of pharmacological induction of RCCs from endogenous CMs, which could offer a promising regenerative strategy to repair injured hearts.

Introduction

Lower organisms like zebrafish exhibit a remarkable capacity for cardiac regeneration through CM dedifferentiation, characterized by the reactivation of embryonic cardiogenic genes and disassembly of sarcomeric structures (Kikuchi & Poss, 2010; Lepilina et al., 2006). In contrast, mammals show limited cardiac regeneration, typically restricted to the early postnatal period (PN1 to PN6) in mice, beyond which the regenerative ability sharply declines (Porrello et al., 2011). This stark difference underscores the need for novel strategies to induce a regenerative state in adult mammalian CMs, akin to that observed in zebrafish, to overcome the inherent barriers to CM dedifferentiation.

During mammalian heart development, CMs arise predominantly from the first heart field (FHF) and the second heart field (SHF) (Srivastava, 2006), with SHF cells marked by the pioneer transcription factor ISL1 proliferating and differentiating into major cardiovascular cell types, CMs, SMCs and ECs during embryonic development (Bu et al., 2009; Moretti et al., 2006). Notably, ISL1+ cells in neonatal mouse hearts show potential for expansion and differentiation into CMs under conducive in vitro conditions (Laugwitz, 2005). Given that ISL1 transcriptionally governs the expression of a suite of cardiac genes, such as NKX2-5 (Ma et al., 2008), FGF10 (Watanabe et al., 2012), which are essential for embryonic cardiogenesis, inducing ISL1 expression in CMs could be a critical determinant in the induction of cells possessing regenerative capacity.

Aiming to replicate this regenerative capacity, our study focused on identifying small molecules capable of inducing ISL1 expression in CMs. Through a combinatorial screening approach, we discovered a novel combination of CHIR99021 and A-485, or 2C, that efficiently and unprecedentedly induces dedifferentiation of human CMs into RCCs. These RCCs exhibited disassembled sarcomeric structures, high expression of embryonic cardiogenic genes, and an increased number of CMs through re-differentiation in vitro. Further investigations showed that 2C robustly generates RCCs in adult mouse hearts and improves cardiac function in mice experiencing myocardial infarction. This proof-of-concept discovery demonstrates that a simple combination of small molecule drugs can endow the mammalian heart with regenerative capacity by reprogramming CMs into RCCs.

Results

Efficient induction of dedifferentiation in hESC-derived CMs by 2C treatment

To obtain an adequate number of CMs for small molecule screening, we differentiated hESCs into CMs (Movie 1, Figure 1—figure supplement 1A) following a well-established step-wise protocol (Lian et al., 2013). Nearly homogeneous contracting CMs were observed on day 10 of differentiation, consistent with previous reports. After purification with glucose-depleted medium containing abundant lactate (Figure 1— figure supplement 1B), highly pure TNNT2+ CMs were obtained. These were subsequently dissociated and seeded into 96-well plates. Once the CMs resumed contraction, they were treated with individual small molecules from a collection of over 4,000 compounds for 3 days (Figure 1—figure supplement 1C), and then fixed and immunostained for ISL1. Using a high-content imaging (HCI) and analysis system, five compounds were initially identified as potential inducers of dedifferentiation in CMs, indicated by induced sarcomere disassembly and ISL1 expression (Figure 1—figure supplement 1D). Further comparison revealed that the unique combination of CHIR99021 and A-485 (2C) most efficiently induced ISL1 expression in TNNT2+ CMs (Figure 1—figure supplement 1E). Notably, while A-485 was screened at a concentration of 10 µM, its optimal working concentration in subsequent experiments was determined to be 0.5 µM based on titration experiments (Figure 1—figure supplement 1F-G). Of note, compound I-BET-762 also showed some capacity to induce ISL1 expression. However, it was less effective than A-485 in combination with CHIR99021 (Figure 1—figure supplement 1H).

2C treatment induced dedifferentiation of hESC-derived CMs toward ISL1-expressing cells.

Cells induced from hESC-derived CMs by treatment with DMSO (NC) or 2C for 60 hours. (A) Phase contrast images showing cell morphology. (B) Immunofluorescence staining of the ISL1 (ISL1, green), and the CM marker cardiac troponin T (TNNT2, red) in the cells. Nuclei were stained by DAPI (4′,6-diamidino-2-phenylindole) and presented in DNA blue. (C), Cytosolic and nuclear areas of the cells. Data are shown as mean ± SD. (n=3 independent experiments, represented as dots). Two-way ANOVA with Dunnett’s multiple comparisons test. ****P < 0.0001. (D) Changes in cell number after 2C treatment. Total cell number, TNNT2+ cell number and ISL1+ cell number were normalized to the negative control DMSO (NC). Data are shown as mean ± SD (n=3 independent experiments, represented as dots). Two-way ANOVA with Šidák’s multiple comparisons test. ns, not significant (P > 0.05), ****P < 0.0001. (E) Fraction of ISL1+ cells in TNNT2+ cells. Data are shown as mean ± SD (n=3 independent experiments, represented as dots). Two-way ANOVA with Šidák’s multiple comparisons test. **P < 0.01. (F) Western blot and quantitative analysis of ISL1 expression in DMSO (NC) or 2C-treated CMs for 60 hours. Data are shown as mean ± SD. Unpaired t test. *P < 0.05. (G) Relative gene expression of embryonic cardiogenesis marker genes (MESP1, ISL1, NR2F2, FUT4, and LEF1), pan-cardiac genes (GATA4, TBX5, and NXK2-5), and CM marker genes (MEF2C, TNNT2, MYL2, and MYL7) in the cells treated by DMSO (NC) or 2C for 60 hours (60h). Data are shown as mean ± SD (n=3 independent experiments, represented as dots). Multiple unpaired t tests. ns, not significant (P > 0.05), *P < 0.05, **P < 0.01, ***P < 0.001, ****P < 0.0001.

When CMs with well-organized sarcomeres were treated with 2C in vitro, cells started to exhibit a dedifferentiation-associated phenotype, such as reduction of cell size after 24 hours, growing as clusters after 48 hours, and forming colonies after 60 hours (Figure 1A and Figure 1—figure supplement 2A). During this period, the expression of TNNT2 and MYL2 gradually downregulated while ISL1 expression and the percentage of ISL1+ cells increased (Figure 1—figure supplement 2B). Compared to untreated CMs, significant sarcomere disassembly and reduction of cytoplasmic/nuclear area were observed in cells treated with 2C for 60 hours (Figure 1B-C). In parallel, the number and percentage of TNNT2+ cells remained constant while ISL1+ cells increased remarkably (Figure 1D-E and Figure 1—figure supplement 2C), accompanied by a ∼3-fold increase in ISL1 expression at both mRNA and protein levels (Figure 1F and Figure 1—figure supplement 2B). Other early embryonic cardiogenesis genes, including MESP1, LEF1 (Klaus et al., 2007), FUT4 (Wang et al., 2019), and NR2F2 (Churko et al., 2018), were also highly expressed following 2C treatment (Figure 1G and Figure 1—figure supplement 2D). Furthermore, 2C treatment significantly decreased the expression of CM-specific genes, such as TNNT2, MYL2, MYL7, and MEF2C, while pan-cardiac transcription factors GATA4, TBX5, and NKX2-5 showed less effect (Figure 1G and Figure 1—figure supplement 2D). Notably, 2C induced ISL1 expression even in mature CMs derived from hESCs treated with ZLN005 (Liu et al., 2020) (Figure 1—figure supplement 2E). These results indicate that 2C treatment effectively reprograms CMs towards a dedifferentiated, embryonic-like state.

2C induced cells possess regenerative capacity

To confirm the regenerative capacity of 2C-induced cardiac cells, we investigated cell proliferation using a BrdU incorporation assay. Immunostaining showed the co-localization of ISL1 and BrdU labeling in these cells (Figure 2A). Statistical analysis revealed a significant decrease in the nuclear area and an increase in the number of ISL1+/BrdU+ positive cells following 2C treatment (Figure 2B-C). We then examined the potential for re-differentiation of 2C-induced cardiac cells into CMs. As expected, upon withdrawal of 2C and subsequent culture in CM media for 3 days (60h+3d), we observed the emergence of spontaneously contracting cells exhibiting CM-specific morphology and typical cytoplasmic/nuclear area size (Movie 2 and Figure 2D-E). These cells downregulated early embryonic cardiogenesis genes and upregulated CM-specific genes, with no noticeable changes in pan-cardiac genes (Figure 2—figure supplement 1A). Remarkably, there was a notable increase (roughly 1.4-fold) in the number of re-differentiated CMs with clear TNNT2 staining relative to the initial CMs before 2C treatment (Figure 2F), demonstrating the regenerative potential in 2C-induced cardiac cells. Additionally, 2C-induced cardiac cells could differentiate into SMA+ SMCs or CD31+/VE-Cadherin+ ECs in the presence of PDGF-BB and TGF-β1 or VEGF, bFGF and BMP4, respectively (Figure 2G-H and Figure 2—figure supplement 1B). In contrast, these genes expressed were not detectable in DMSO-treated cells. Collectively, our findings indicate that 2C treatment facilitates the conversion of CMs into cardiac cells with regenerative capability, including proliferative potential and differentiation into the three major cardiovascular cell types, thus named regenerative cardiac cells (RCCs).

Regenerative ability of 2C-induced RCCs.

(A-C) Immunofluorescence staining (A) and statistical analysis (B and C) of the ISL1 (green) and BrdU (red) double positive RCCs induced from CMs by treatment with DMSO (NC) or 2C for 60 hours. The ISL1+ cell number and ISL1+/BrdU+ cell number were normalized to the negative control DMSO (NC). DAPI (4′,6-diamidino-2-phenylindole) staining labeled nuclei as blue. Data are shown as mean ± SD (n=3 independent experiments, represented as dots). Multiple unpaired t tests in (B), ****P < 0.0001. Two-way ANOVA with Šidák’s multiple comparisons test in (C), **P < 0.01, ****P < 0.0001. (D) Phase contrast images of hESC-derived CMs treated by DMSO (NC) or 2C for 60 hours (60h) and subsequently cultured in the absence of 2C for another 3 days (60h+3d). (E) Immunostaining showed the expression of ISL1 (green) and TNNT2 (red) in the cells under the same condition in (D). (F) Statistical analysis of TNNT2+ cell numbers under the same condition in (D). Data are shown as mean ± SD (n=3 independent experiments, represented as dots). Two-way ANOVA with Šidák’s multiple comparisons test. ns, not significant (P > 0.05), ****P < 0.0001. (G) Schematic diagram of directed differentiation of 2C-induced RCCs towards cardiomyocytes (CMs), smooth muscle cells (SMCs) and endothelial cells (ECs). (H) Immunostaining showed the expression of EC markers (CD31, green and CD144, red), SMC marker (SMA, red), and CM marker (TNNT2, green). DAPI (4′,6-diamidino-2-phenylindole) staining labeled nuclei as blue.

Lineage tracing demonstrated that 2C induced RCCs dedifferentiated from TNNT2+ CMs

Although TNNT2+ CMs purified using a lactate-based culture medium were nearly homogeneous populations, a small fraction (<4%) of purified cells still expressing ISL1 (Figure 1—figure supplement 1B). To ascertain that the RCCs induced by 2C treatment were indeed dedifferentiated from TNNT2+ CMs rather than originating from the proliferation of residual ISL1+ cells, we conducted lineage-tracing experiments. Using an ISL1mCherry/+ H9 hESC (K9) line established through CRISPR-based knock-in (Figure 3—figure supplement 1A-D), we verified that mCherry expression accurately mirrored endogenous ISL1 expression (Figure 3A-C). Following the purification of K9-derived CMs and the selection of mCherry-negative cells by FACS (Figure 3D-E), these cells were confirmed to be TNNT2+ CMs (Figure 3—figure supplement 1E). Upon treatment of these mCherry-negative CMs with 2C for 60 hours, we observed remarkable morphological changes in mCherry+/ISL1+ RCCs compared to mCherry−/TNNT2+ CMs (Figure 3F), accompanied by a dramatic downregulation of TNNT2 and MYL2 expression and a notable upregulation of ISL1 and mCherry expression (Figure 3G). FACS analysis further validated that 2C could convert K9-derived mCherry-negative CMs into mCherry+/ISL1+ cells (Figure 3H-I). Similar findings were obtained using the HUES7 hESC line with an ISL1-mCherry knock-in reporter (K7) (Figure 3—figure supplement 2), robustly demonstrating that 2C induces ISL1 expression in TNNT2+ CMs, effectively generating RCCs.

Lineage tracing demonstrated 2C induced dedifferentiation of TNNT2+ CMs to ISL1-expressing RCCs.

(A) Immunofluorescence images showing expression of endogenous ISL1 (green) and ISL1-mCherry (red) reporter in the cells differentiated from K9 hESC KI reporter line at day 6 (D6). DAPI (4′,6-diamidino-2-phenylindole) staining labeled nuclei as blue. (B-C) Flow cytometry analysis of the percentage of mCherry+/ISL1+ cells in the cells differentiated from K9 at D6. (D-E) Flow cytometry analysis of the percentage of mCherry-negative cells at selection day 4 (SD4) in lactate purification medium. (F) Cells induced from mCherry-negative CMs by treatment with or without 2C for 60 hours. Images showing the expression of mCHERRY (red) and TNNT2 (green) in the cells. (G) Relative gene expression of ISL1, mCHERRY, LEF1, TNNT2 and MYL2 in K9-derived mCherry-negative CMs treated with DMSO (NC) or 2C for 60 hours. Data are shown as mean ± SD (n=2 independent experiments with 4 replicates each). Two-way ANOVA with Šidák’s multiple comparisons test. ***P < 0.001, ****P < 0.0001. (H-I) Flow cytometry analysis of the percentage of mCherry-positive cells induced from mCherry-negative CMs by treatment with or without 2C for 60 hours. Data are shown as mean ± SD (n=3 independent experiments, represented as dots). (J-K) Flow cytometric plots showing EGFP-labeled CMs by lineage-tracing of K9-derived mCherry-negative CMs (J), and bar graph showing the percentage of mCherry-negative CMs expressing EGFP (K). Data are shown as mean ± SD (n=3 independent experiments). (L) Images showing the expression of ISL1 (red) and EGFP (green) in the cells induced from EGFP-positive/mCherry-negative CMs by treatment with or without 2C for 60 hours. DAPI (4′,6-diamidino-2-phenylindole) staining labeled nuclei as blue.

Additionally, we used a lineage-tracing system to unequivocally confirm the conversion of CMs to ISL1-expressing RCCs by 2C treatment. In this assay, CMs were transfected with plasmids encoding CreERT2 under the control of the TNNT2 promoter and an EGFP reporter following a flox-stop-flox cassette (Figure 3—figure supplement 3A). After optimizing viral titers and infection timing (Figure 3—figure supplement 3B-D), approximately 0.6% of K9-derived mCherry-negative CMs were permanently labeled with EGFP following 6 days of tamoxifen treatment (Figure 3J-K). These EGFP-labeled mCherry-negative CMs were then sorted by flow cytometry and treated with 2C for 60 hours. As expected, ISL1 expression was observed in these EGFP+ CMs (Figure 3L). Collectively, these results strongly indicate that the RCCs with ISL1 expression were genuinely generated by 2C treatment from TNNT2+ CMs.

2C-induced dedifferentiation of CMs provided a protective effect on cardiac infarction

To examine the reprogramming effect of 2C on CMs in vivo, we first tested 2C on primary neonatal rat CMs (Sakurai et al., 2014), which consistently induced RCC generation with corresponding ISL1 induction and morphological changes (Figure 4—figure supplement 1A). We then evaluated whether 2C could reprogram endogenous CMs to RCCs in vivo. Neonatal SD rats were intraperitoneally administrated 20 mg/kg of CHIR99021 and 10 mg/kg of A-485 daily for 5 days (Figure 4—figure supplement 1B). Compared to the vehicle/DMSO controls (NC), ISL1 was robustly induced in the CMs of rats treated with 2C without apparent effects on heart to body weight ratio (Figure 4—figure supplement 1C-D). These induced ISL1-expressing cells were widely distributed within the region 600 µm to 3000 µm down from the base of the hearts, including the aorta (Ao), left/right atria, and both ventricles (Figure 4—figure supplement 1E). All these ISL1+ cells also expressed TNNT2, indicating that 2C-induced RCCs originated from CMs in vivo. Importantly, neither CHIR99021 nor A-485 alone induced ISL1+ cells in ventricles, highlighting the combined effect of 2C on RCC induction (Figure 4—figure supplement 2).

Heart regeneration via 2C-induced dedifferentiation of CMs.

(A) Immunofluorescence staining of ISL1 (green) and TNNT2 (red) in cross-sectioned hearts from 2C or vehicle (DMSO)-treated (NC) adult 129SvJ mice. Ao, aorta. PA, pulmonary artery. LA, left atrial. RA, right atrial. IAS, interatrial septum. DAPI (4′,6-diamidino-2-phenylindole) staining labeled nuclei as blue. (B) Schematic illustration of the method used to examine the prophylactic effect of 2C in 129SvJ mice post MI. (C) Body weight of mice pre-treated with vehicle (DMSO) or 2C as shown in (B) at Day 12, Day 16, Day 20, and Day 35 after MI (dpi). Error bars represent SD. ns, not significant (P > 0.05). (D) Survival curve of sham-operated mice and mice pre-treated with vehicle (DMSO) or 2C as shown in (B), at indicated time points before or after MI. (E) Ejection fraction (EF) of sham-operated mice and mice pre-treated with vehicle (DMSO) or 2C as shown in (B), before MI (baseline) or at Day 1, Day 8, Day 25, and Day 35 after MI (dpi). Data are shown as mean ± SD. Two-way ANOVA with Tukey’s multiple comparisons test. ns, not significant (P > 0.05), **P < 0.01, ***P < 0.001. (F) Schematic illustration of the method used to examine therapeutic effect of 2C in the 129SvJ mice post MI. (G) Serial fMRI measurements showing the cardiac function from sham-operated mice and mice treated with vehicle (DMSO) or 2C at as shown in (F). Data are shown as mean ± SD. Two-way ANOVA with Tukey’s multiple comparisons test. *P < 0.05. (H) Echocardiography of sham-operated mice and mice treated with vehicle (DMSO) or 2C as shown in (F) at Day 35 post MI. (I) Masson staining of serial transverse sections of hearts from sham-operated mice and mice treated with vehicle (DMSO) or 2C as shown in (F) at Day 35 post MI.

Similarly, RCCs were efficiently induced in the Ao root and RA regions of the heart dissected from adult 129SvJ mice administrated 2C (20 mg/kg of CHIR99021 and 10 mg/kg of A-485) for 5 consecutive days (Figure 4A). To explore whether 2C-induced RCCs could rescue cardiac function in mice subjected to myocardial infarction (MI), we conducted both prophylactic and therapeutic studies. In the prophylactic setting, mice were pre-treated with either 2C or vehicle/DMSO for 5 days (Figure 4B) before inducing MI by ligation of left anterior descending (LAD) artery. When measuring cardiac function with magnetic resonance imaging (MRI) on days 1, 8, 25 and 35 post-MI, we found pretreatment with 2C significantly improved both cardiac function and survival rate without affecting body weight (Figure 4C-E). In the therapeutic setting (Figure 4F), mice treated with 2C post-MI displayed recovered cardiac function as assessed by left ventricular ejection fraction (LVEF) and echocardiography (Figure 4G-H). Cardiac fibrosis was largely ameliorated in 2C-treated mice, displaying significantly smaller scar sizes compared to control mice (Figure 4I). Altogether, these in vivo studies collectively indicate that 2C-induced RCCs possess regenerative capacities comparable to those generated in vitro, shedding light on the development of small molecule drugs with similar mechanisms for therapeutic use in repairing or regenerating hearts after cardiac injury.

Dedifferentiation requirement of CHIR99021 and A-485

To elucidate the mechanisms behind the in 2C-induced dedifferentiation of CMs into RCCs, we performed bulk RNA-seq on K9-derived ISL1/mCherry-negative CMs treated with either DMSO as a negative control (NC) or the 2C combination for 60 hours (Figure 5A). Analysis of differentially expressed genes (DEGs) revealed substantial transcriptomic alterations; specifically, 2C treatment upregulated embryonic cardiogenesis genes such as MSX1, BMP4, TCF4, and LEF1, and downregulated genes associated with CMs (Figure 5B-C). Gene ontology (GO) analysis further indicated a suppression of genes involved in cardiac maturation and muscle contraction, contrasting with an enrichment in genes part of the catenin complex (Figure 5D-F). These findings were corroborated by qPCR, confirming the significant up-regulation of embryonic genes (e.g., MSX2, NKD1, PDGFC, CTNNA2) in K9-derived mCherry-negative CMs (Figure 5G).

Bulk RNA-seq of analysis of 2C-treated ISL1/mCherry-negative CMs.

(A) Scheme of bulk RNA-seq analysis of K9-derived mCherry-negative CMs with DMSO (NC) or 2C treatment for 60 hours. (B) Heatmap of differentially expressed genes (DEGs) in ISL1/mCherry-negative CMs treated with DMSO (NC) or 2C for 60 hours. (C) Volcano plot showing genes significantly changed by 60 hours of 2C treatment. (D-E) Gene Ontology (GO) analysis of downregulated (D) and upregulated (E) genes in ISL1/mCherry-negative CMs by 2C treatment for 60 hours, compared to DMSO (NC) treated cells. (F) Plotting GO terms of upregulated genes by 2C treatment with cnetlpot. (G) Relative expression fold-changes of indicated genes in K9-derived ISL1/mCherry-negative CMs by 60 hours of DMSO (NC) or 2C treatment. Data are shown as mean ± SD. Multiple unpaired t tests. **P < 0.01, ***P < 0.001.

We investigated the individual and combined effects of CHIR99021 and A-485 on inducing RCC characteristics. Morphological assessments showed that CHIR99021 significantly reduced both cytoplasmic and nuclear areas—a feature characteristic of RCCs—more effectively than A-485, which primarily reduced cytoplasmic area only (Figure 5—figure supplement 1A-C). Sarcomere disassembly was more pronounced with CHIR99021 treatment, aligning with its role in enhancing embryonic cardiogenesis genes like BMP4 and LEF1 (Figure 5—figure supplement 1D-E). A-485, while not affecting the nuclear area to the same extent, played a crucial role in the epigenetic reprogramming of CMs. It facilitated the suppression of CM-specific genes, such as TNNT2, through epigenetic modifications, enhancing the expression of RCC-specific genes such as ISL1 in conjunction with CHIR99021 (Figure 5—figure supplement 1D). This indicates that while CHIR99021 drives the initial dedifferentiation process by altering the cellular morphology and gene expression associated with embryonic cardiogenesis, A-485 enhances these effects by providing necessary epigenetic conditions for sustaining and stabilizing the RCC phenotype. Despite both compounds independently inducing reduced expression of TNNT2 and MYL2, as well as increased ISL1 expression, CHIR99021 uniquely influences a cohort involved in embryonic cardiogenesis, including BMP4, NKD1, MSX2, PDGFC, LEF1, and TCF4 (Figure 5—figure supplement 1D). Upon withdrawal of CHIR99021, A-485 or 2C, the cell percentage of ISL1+ cell was similarly and significantly reduced (Figure 5—figure supplement 1F). Notably, a drastic increase in the number of TNNT2+ CMs was observed only during the redifferentiation of 2C-treated cells, as confirmed by statistical analysis (Figure 5—figure supplement 1G). These findings suggest that while CHIR99021 plays a leading role in 2C-induced de-differentiation of CMs to RCCs, A-485 is also indispensable, particularly in modifying the epigenetic landscape to support the transition and maintenance of reprogrammed state.

Reprogramming of CMs to RCCs by 2C went through an intermediate cell state

In the exploration of CM reprogramming to RCC using 2C treatment, single-cell RNA sequencing (scRNA-seq) was utilized to trace the progression through distinct cellular states of K9-derived mCherry-negative CMs treated with either DMSO (NC) or the 2C for 60 hours. Through UMAP analysis, we identified seven distinct clusters, observing notable increases in the proportions of cells in clusters 0, 2 and 3 following 2C treatment (Figure 6A-B). Cells within cluster 2 prominently expressed genes characteristic of RCCs, such as ISL1, BMP4 and FGF20—a gene crucial for the expansion of early embryonic progenitor cells (Cohen et al., 2007)), along with the cell proliferation marker MKI67 (Figure 6C-D). This distinct expression pattern marked them as transitioning towards an RCC phenotype. In contrast, clusters 0, 3, 4, and 5 maintained high expression levels of CM-specific markers like MYH6 and MYL2 (Figure 6C-D), indicating their retention of a mature CM identity. Interestingly, cells in clusters 1 and 6 exhibited characteristics of intermediate cells (ICs), expressing both dedifferentiation markers such as ACTA2 and genes linked to CM development, including COL1A1, COL1A2, and COL3A1 (Cui et al., 2020; Mononen et al., 2020) (Figure 6C-D). This highlighted these clusters as transitional states in the reprogramming process from CMs to RCCs.

Single-cell RNA-seq of 2C-treated mCherry-negative CMs.

(A) UMAP analysis showing 7 clusters in cells induced from K9-derived mCherry-negative CMs by treatment with DMSO (NC) and 2C for 60 hours. (B) The percentage of cells in the 7 indicated clusters, following DMSO (NC) or 2C treatment. (C) Heatmap showing the differentially expressed genes in the cells from 7 indicated clusters. The representative marker genes of 7 indicated clusters were listed on the right. (D) Violin plots showing the expression levels of marker genes of CMs (MYL2, MYH6), ICs (COL1A1, ACTA2), and RCCs (ISL1, BMP4, FGF20) among cells from 7 indicated clusters. (E) UMAP analysis showing the second-level clustering of cluster 2 into 4 subclusters (left), which exhibited dramatic distinction under condition of 2C or NC (right). (F) The percentage of cells in the 4 indicated subclusters within cluster 2, following DMSO (NC) or 2C treatment. (G) Heatmap showing the differentially expressed genes among cells from 4 subclusters of cluster 2. Genes related to RCCs are highlighted in the green blocks on the right. (H-I) Pseudotime trajectory showing changes across various cell states upon 2C treatment, which were presented with different developmental pseudotime points (H) and cell states (I). (J) Curves showing the dynamic expression of representative genes of RCCs (ISL1, FGF20), ICs (COL1A1), and CMs (TNNT2) along indicated pseudotime points.

Further detailed analysis of cluster 2, comparing 435 cells from DMSO-treated samples with 410 cells from 2C-treated samples, allowed us to identify a prominent subcluster (subcluster 0) comprising 408 cells (48.3% of the cluster), exclusively found in 2C-treated samples (Figure 6E-F). Cells in this subcluster were defined by high expression of key embryonic cardiogenesis genes such as MSX1, LEF1, BMP4, MSX2, and HAND1, along with genes typically co-expressed with ISL1 in SHF progenitors, including NR2F2, TBX5, ALCAM (Ghazizadeh et al., 2018), and CXCR4 (Andersen et al., 2018) (Figure 6G). This specific gene expression profile robustly indicates that the cells in subcluster 0 have adopted an RCC identity, indicative of successful induction by 2C treatment. Thus, a unique gene set that included LIX1, NKD1, PDGFC, ARL4A, AXIN2, FZD7, ISL1, MSX1, MSX2, BMP4, LEF1, and HAND1 was found to determine RCC state during the reprogramming of CMs by 2C, and hence designated as RCC genes.

To further quantify this transition, pseudo-time analysis was performed, clearly delineating the 2C-induced reprogramming trajectory from CMs, highly expressing TNNT2, through ICs, characterized by the expression of COL1A1, to RCCs, marked by elevated levels of ISL1 and FGF20) (Figure 6H-J). This analysis confirms the dynamic and stepwise nature of cellular transformation under 2C treatment.

Epigenetic regulation of CMs and RCCs genes by 2C treatment

As previously documented (Lasko et al., 2017), A-485 functions as a p300 acetyltransferase inhibitor, specifically reducing acetylation at lysine 27 of histone H3 (H3K27Ac), but not at lysine 9 (H3K9Ac), as shown in studies with PC3 cells. Our data corroborate these findings, showing a notable decrease in H3K27Ac levels in CMs treated with A-485 alone, while levels of H3K9Ac remain unaffected (Figure 7— figure supplement 1A-E). However, treatment with CHIR99021 or the combined 2C regime did not significantly alter the acetylation patterns of these histone marks (Figure 7—figure supplement 1). Intriguingly, we observed that although the number of H3K9Ac+ cells remained unchanged in CMs, the fluorescence intensity was significantly decreased after A-485 treatment (Figure 7—figure supplement 1F), suggesting a distinctive regulatory role of A-485 in modulating histone acetylation in CM reprogramming.

Chromatin immunoprecipitation-sequencing (ChIP-seq) analyses of chemical-treated H9 hESC-derived CMs.

(A) Schematic illustration of ChIP-seq analysis of H9-derived CMs subjected to DMSO, A-485, CHIR99021, or 2C treatment for 60 hours. (B) Average ChIP-seq signal profiles showing the indicated histone modifications around the TSS in the input and ChIP samples prepared from DMSO and 2C-treated cells. (C) Heatmap showing the whole-genome wide distribution of H3K9Ac, H3K27Ac, and H3K4me3 peaks within a range of ±4kb from TSSs in the cells treated with DMSO or 2C for 60 hours. (D) H3K9Ac and H3K27Ac peaks surrounding CM genes (TNNT2 and TNNI1) and RCC genes (LEF1 and ISL1) in the cells treated with DMSO, A-485, CHIR99021 or 2C for 60 hours, and H3K4me3 peaks surrounding the same genes in the cells treated by DMSO or 2C. The y-axis represents the number of counts. (E-F) Veen diagram showing the number of annotated genes with H3K9Ac or H3K27Ac enrichment in the cells treated with DMSO, A-485, CHIR99021 or 2C for 60 hours. Red circles indicate the number of genes with unique H3K9Ac or H3K27Ac enrichment induced by 2C treatment; black circles indicate the number of genes with H3K9Ac or H3K27Ac enrichment unaffected by any chemical treatment. (G-J) The annotated genes with the most significant changes in H3K9Ac enrichment following treatment with DMSO (G) or 2C (I) were ranked by Log10LR and analyzed by GOs (H and J), respectively. (K) ISL1 binding motifs identified from the cells treated with A-485 or 2C.

To explore the molecular mechanisms by which the 2C treatment modulates gene expression during the reprogramming of hESC-derived CMs into RCCs, we conducted chromatin immunoprecipitation followed by sequencing (ChIP-seq) analyses. This study focused on H3K27Ac and H3K9Ac modifications across various treatment conditions (Figure 7A). Our analysis identified regions near transcription start sites (±4 kb) that were rich in H3K4me3, indicative of active promoters in both control (DMSO-treated) and 2C-treated cells (Figure 7B-C). Following 2C treatment, there was a marked increase in H3K9Ac and H3K27Ac levels around the transcription start sites (TSS), suggesting enhanced gene activity. Specifically, 2C treatment significantly increased H3K9Ac enrichment at 4,485 gene promoters and H3K27Ac at 2,560 gene promoters, with corresponding decreases in 346 and 1,834 genes, respectively (Figure 7—figure supplement 2A-F). Notably, genes that showed reduced acetylation in response to A-485 also displayed similar trends under 2C treatment, highlighting genes with functions crucial to cardiac muscle activity such as contraction and sarcomere organization (Figure 7—figure supplement 2). In comparison with treatments using A-485 or CHIR99021 alone, the 2C combination notably augmented acetylation at genes involved in cell cycle regulation and cell division, suggesting a synergistic effect of the dual treatment in promoting gene activation essential for cardiac regeneration (Figure 7—figure supplement 2).

We also specifically analyzed the acetylation peaks at promoters of key CM genes and RCC genes. A-485 treatment predominantly decreased H3K9Ac and H3K27Ac peaks at CM gene promoters such as TNNT2, TNNI1, MYL7, MYH6, and MYH7, aiding their downregulation. These reductions were intensified under 2C treatment, reflecting its potent effect in repressing mature cardiomyocyte markers while activating regenerative pathways (Figure 7D and Figure 7—figure supplement 3). In contrast, CHIR99021 effectively upregulated H3K9Ac and H3K27Ac at promoters of the RCC genes such as LEF1, AXIN2, BMP4, LIX1, MSX1, MSX2, and NKD1, with 2C treatment further enhancing these effects, confirming its robust impact on inducing a regenerative phenotype in treated cells (Figure 7D and Figure 7—figure supplement 3).

Moreover, the distribution of H3K9Ac and H3K27Ac changes during 2C-induced dedifferentiation revealed that these modifications are more dynamically associated with H3K9Ac, particularly in the transition of CMs to RCCS. This dynamic regulation was illustrated by the significant number of genes showing exclusive enrichment under 2C treatment compared to control or single-drug treatments (Figure 7E-F). In detail, of genes with H3K9Ac enrichment, 2716 (or 46.1%) out of 5891 annotated genes were only enriched in the cells treated with 2C, while only 527 (or 4.9%) out of 10785 annotated genes with H3K27Ac enrichment were exclusively observed (Figure 7E-F). To further confirm this notion, we performed a comparative analysis of annotated genes between the cells treated with 2C or DMSO to validate the most significant differential genes (i.e., Log10LR>3) under the respective condition. In comparison to 430 genes with H3K9Ac enrichment observed in DMSO-treated cells, which were mainly involved in cardiac muscle contraction (Figure 7G-H), 2C treatment markedly enhanced the enrichment of H3K9Ac in 123 genes associated with the transcription factor complex and Wnt signaling pathway (Figure 7I-J). In particular, 12 CM genes, including TNNT2, were ranked in the top 90 out of 430 annotated genes with H3K9Ac enrichment. In contrast, 7 RCC genes, including ISL1, were ranked in the top 50 out of 123 annotated genes with H3K9Ac enrichment.

Lastly, our findings were substantiated by DNA-binding motif analyses, which showed increased accessibility of ISL1 binding sites only in cells treated with A-485 or the 2C combination, underscoring the critical role of A-485 in activating RCCs genes through ISL1-dependent mechanisms (Figure 7K). Collectively, these results underscore the efficiency of combining CHIR99021 and A-485 in reprogramming CMs to RCCs, opening new avenues for developing therapies aimed at cardiac regeneration.

Discussion

The generation of regenerative cardiac cells (RCCs) from cardiomyocytes (CMs) using the 2C treatment, a combination of CHIR99021 and A-485, represents a significant advancement in cardiac regenerative medicine. This study demonstrates that the synergistic interaction between these two molecules is essential for the reprogramming of endogenous CMs into RCCs, as evidenced by the absence of RCCs in postnatal rat hearts treated with either molecule alone (Figure 4—figure supplement 2). CHIR99021, a well-documented GSK3 inhibitor, activates the Wnt signaling pathway, thereby promoting the proliferation of human pluripotent stem cell-derived CMs (Buikema et al., 2020; Quaife-Ryan et al., 2020). It also induces H3K27Ac at promoters of critical cardiac genes such as TNNT2, MYH6, and ISL1 (Quaife-Ryan et al., 2020). This observation aligns with our findings, which show a slight increase H3K27Ac levels in CMs following CHIR99021 treatment (Figure 7D and Figure 7—figure supplement 1). Moreover, our results reveal that the p300/CBP inhibitor A-485 facilitates CM dedifferentiation into RCCs by epigenetically suppressing CM-specific gene expression (Figure 7D), consistent with a decline in the fluorescence intensity of H3K27Ac and H3K9Ac in CMs (Figure 7—figure supplement 1C and E). Concurrently A-485 enhances modifications conducive to the RCC state, such as H3K9Ac and H3K27Ac on genes like ISL1 (Figure 7D). This dual mechanism of action underscores the potential of utilizing epigenetic modifiers like A-485 to influence cell fate decisions in terminally differentiated cells, suggesting broader applications in regenerative biology.

The RCCs induced by the 2C treatment not only re-expressed genes essential for embryonic cardiogenesis but also demonstrated the capability to differentiate into three cardiac lineages in vitro, mirroring the properties of ISL1+ second heart field cardiac progenitors (Bu et al., 2009; Moretti et al., 2006). Although these 2C-induced RCCs show limited proliferation compared to natural ISL1+ progenitors, their potential for expansion under defined conditions presents an intriguing avenue for future research. These insights contribute significantly to the understanding of cardiac lineage plasticity and highlight the therapeutic potential of induced multipotent cardiovascular progenitors for heart repair and regeneration.

Despite these promising findings, several limitations warrant further investigation. The modest proliferation ability of 2C-induced RCCs relative to natural cardiac progenitors suggests that optimizing the culture conditions or treatment protocols may enhance their regenerative capacity. Additionally, the specific molecular mechanisms through which CHIR99021 and A-485 synergistically promote RCC formation remain incompletely understood. Future studies should aim to delineate these pathways more clearly, potentially through in vivo lineage tracing to map the fate decisions of individual cells. Moreover, the potential off-target effects of 2C treatment and its long-term impacts on cardiac function and structure need comprehensive evaluation. As we advance, exploring the scalability of this approach and its applicability to other types of terminally differentiated cells could broaden the scope of regenerative therapies, offering new strategies for numerous degenerative conditions. In conclusion, this study not only sheds light on the cellular and molecular intricacies of cardiac cell dedifferentiation but also opens new pathways for the development of regenerative therapies aimed at heart disease.